### 2020-07-12 cotyledon-data-analysis.Rmd
data analysis of cotyledons

---
title: "WI FastPlants: Heritability of number of cotelydons"
author: "Claudia Solis-Lemus"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(car)
```

### Reading the data

Reading in the data (had to remove manually the '#' sign from the column and save as csv file):
```{r}
dat = read.table("data/Polycot-data-from-NY.csv", sep=",",header=TRUE)
str(dat)
```

The data has 4 columns:
```{r}
head(dat)
dat[200:210,]
```

We can summarize basic counts. For example, we have 622 offspring plants and 181 parent plants. On average, the plants have 2.691 cotelydons (`mean=2.691`). The median being 3 cotelydons means that 50% of the plants have 3 cotelydons or fewer
```{r}
summary(dat)
summary(dat$parents)
```

# Data visualization

```{r, echo=FALSE}
p1 = ggplot(dat, aes(generation, fill=generation))+geom_bar()+
  theme(
        plot.title = element_text(hjust=0.5, size=rel(1.8)),
        axis.title.x = element_text(size=rel(1.8)),
        axis.title.y = element_text(size=rel(1.8), angle=90, vjust=0.5, hjust=0.5),
        axis.text.x = element_text(colour="grey", size=rel(1.5), angle=0, hjust=.5, vjust=.5, face="plain"),
        axis.text.y = element_text(colour="grey", size=rel(1.5), angle=0, hjust=.5, vjust=.5, face="plain"),
        panel.background = element_blank(),
        axis.line = element_line(colour = "grey")##,
        )
p1
```
```{r, echo=FALSE}
p1.2 = ggplot(dat, aes(parents, fill=parents))+geom_bar()+
  theme(
        plot.title = element_text(hjust=0.5, size=rel(1.8)),
        axis.title.x = element_text(size=rel(1.8)),
        axis.title.y = element_text(size=rel(1.8), angle=90, vjust=0.5, hjust=0.5),
        axis.text.x = element_text(colour="grey", size=rel(1.5), angle=0, hjust=.5, vjust=.5, face="plain"),
        axis.text.y = element_text(colour="grey", size=rel(1.5), angle=0, hjust=.5, vjust=.5, face="plain"),
        panel.background = element_blank(),
        axis.line = element_line(colour = "grey")##,
        )
p1.2
```
```{r, echo=FALSE}
p2 = ggplot(dat, aes(x = parents, y = cotyledons, fill = parents)) +
  geom_point(pch = 21, alpha=0.3, position = position_jitterdodge(jitter.height=0.075, jitter.width=2.5))+
  ylim(c(1,6))+
  theme(
        plot.title = element_text(hjust=0.5, size=rel(1.8)),
        axis.title.x = element_text(size=rel(1.8)),
        axis.title.y = element_text(size=rel(1.8), angle=90, vjust=0.5, hjust=0.5),
        axis.text.x = element_text(colour="grey", size=rel(1.5), angle=0, hjust=.5, vjust=.5, face="plain"),
        axis.text.y = element_text(colour="grey", size=rel(1.5), angle=0, hjust=.5, vjust=.5, face="plain"),
        panel.background = element_blank(),
        axis.line = element_line(colour = "grey")##,
        )
p2
```

```{r, echo=FALSE}
p2.2 = ggplot(dat, aes(x=parents, y=cotyledons, fill=parents))+geom_boxplot(alpha=0.4)+
  ylim(c(1,6))+
  theme(
        plot.title = element_text(hjust=0.5, size=rel(1.8)),
        axis.title.x = element_text(size=rel(1.8)),
        axis.title.y = element_text(size=rel(1.8), angle=90, vjust=0.5, hjust=0.5),
        axis.text.x = element_text(colour="grey", size=rel(1.5), angle=0, hjust=.5, vjust=.5, face="plain"),
        axis.text.y = element_text(colour="grey", size=rel(1.5), angle=0, hjust=.5, vjust=.5, face="plain"),
        panel.background = element_blank(),
        axis.line = element_line(colour = "grey")##,
        )
p2.2
```

```{r, echo=FALSE}
p3 = ggplot(dat, aes(x = parents, y = cotyledons, fill = parents)) +
  geom_boxplot(outlier.size = 0, alpha=0.1) +
  geom_point(pch = 21, alpha=0.3, position = position_jitterdodge(jitter.height=0.05, jitter.width=2.5))+
  ylim(c(1,6))+
  theme(
        plot.title = element_text(hjust=0.5, size=rel(1.8)),
        axis.title.x = element_text(size=rel(1.8)),
        axis.title.y = element_text(size=rel(1.8), angle=90, vjust=0.5, hjust=0.5),
        axis.text.x = element_text(colour="grey", size=rel(1.5), angle=0, hjust=.5, vjust=.5, face="plain"),
        axis.text.y = element_text(colour="grey", size=rel(1.5), angle=0, hjust=.5, vjust=.5, face="plain"),
        panel.background = element_blank(),
        axis.line = element_line(colour = "grey")##,
        )
p3
```

```{r, echo=FALSE}
p4 = ggplot(dat, aes(x=parents, y=cotyledons, fill=parents))+geom_violin(alpha=0.5)+
  ylim(c(1,6))+
  theme(
        plot.title = element_text(hjust=0.5, size=rel(1.8)),
        axis.title.x = element_text(size=rel(1.8)),
        axis.title.y = element_text(size=rel(1.8), angle=90, vjust=0.5, hjust=0.5),
        axis.text.x = element_text(colour="grey", size=rel(1.5), angle=0, hjust=.5, vjust=.5, face="plain"),
        axis.text.y = element_text(colour="grey", size=rel(1.5), angle=0, hjust=.5, vjust=.5, face="plain"),
        panel.background = element_blank(),
        axis.line = element_line(colour = "grey")##,
        )

p4
```
```{r, echo=FALSE}
p4.2 = ggplot(dat, aes(x=parents, y=cotyledons, fill=parents))+geom_violin(alpha=0.5)+
    geom_point(pch = 21, alpha=0.3, position = position_jitterdodge(jitter.height=0.05, jitter.width=2.5))+
  ylim(c(1,6))+
  theme(
        plot.title = element_text(hjust=0.5, size=rel(1.8)),
        axis.title.x = element_text(size=rel(1.8)),
        axis.title.y = element_text(size=rel(1.8), angle=90, vjust=0.5, hjust=0.5),
        axis.text.x = element_text(colour="grey", size=rel(1.5), angle=0, hjust=.5, vjust=.5, face="plain"),
        axis.text.y = element_text(colour="grey", size=rel(1.5), angle=0, hjust=.5, vjust=.5, face="plain"),
        panel.background = element_blank(),
        axis.line = element_line(colour = "grey")##,
        )

p4.2
```

Now, we can combine groups in our data:

- Parents
- Offspring with both parents having 2 cotyledons
- Offspring with one parent having more than 2 cotyledons
- Offspring with both parents having more than 2 cotelydons
```{r}
dat2 = within(dat, group <- "(2+)x(2+)")
dat2$group[dat2$parents == ""] = "parent"
dat2$group[dat2$parents == "2x2"] = "2x2"
dat2$group[dat2$parents == "2x3"] = "2x(2+)"
dat2$group[dat2$parents == "2x4"] = "2x(2+)"
dat2$group = factor(dat2$group,levels=c("parent", "2x2", "2x(2+)", "(2+)x(2+)"))
```

```{r, echo=FALSE}
p5 = ggplot(dat2, aes(x=group, y=cotyledons, fill=group))+geom_violin(alpha=0.5, trim=FALSE)+
  ylim(c(1,6))+
  stat_summary(fun.y=mean, geom="point", shape=23, size=2)+
  theme(
        plot.title = element_text(hjust=0.5, size=rel(1.8)),
        axis.title.x = element_text(size=rel(1.8)),
        axis.title.y = element_text(size=rel(1.8), angle=90, vjust=0.5, hjust=0.5),
        axis.text.x = element_text(colour="grey", size=rel(1.5), angle=0, hjust=.5, vjust=.5, face="plain"),
        axis.text.y = element_text(colour="grey", size=rel(1.5), angle=0, hjust=.5, vjust=.5, face="plain"),
        panel.background = element_blank(),
        axis.line = element_line(colour = "grey")##,
        )

p5
```

```{r, echo=FALSE}
p5.2 = ggplot(dat2, aes(x=group, y=cotyledons, fill=group))+geom_violin(alpha=0.5, trim=FALSE)+
  ylim(c(1,6))+
  stat_summary(fun.y=mean, geom="point", shape=23, size=2)+
  geom_point(pch = 21, alpha=0.3, position = position_jitterdodge(jitter.height=0.05, jitter.width=2.5))+
  theme(
        plot.title = element_text(hjust=0.5, size=rel(1.8)),
        axis.title.x = element_text(size=rel(1.8)),
        axis.title.y = element_text(size=rel(1.8), angle=90, vjust=0.5, hjust=0.5),
        axis.text.x = element_text(colour="grey", size=rel(1.5), angle=0, hjust=.5, vjust=.5, face="plain"),
        axis.text.y = element_text(colour="grey", size=rel(1.5), angle=0, hjust=.5, vjust=.5, face="plain"),
        panel.background = element_blank(),
        axis.line = element_line(colour = "grey")##,
        )

p5.2
```

```{r, echo=FALSE}
p5.3 = ggplot(dat2, aes(x=group, y=cotyledons, fill=group))+geom_violin(alpha=0.5, trim=FALSE)+
  ylim(c(1,6))+
  geom_point(pch = 21, alpha=0.3, position = position_jitterdodge(jitter.height=0.05, jitter.width=2.5))+
    stat_summary(fun.y=median, geom="point", size=2)+
  theme(
        plot.title = element_text(hjust=0.5, size=rel(1.8)),
        axis.title.x = element_text(size=rel(1.8)),
        axis.title.y = element_text(size=rel(1.8), angle=90, vjust=0.5, hjust=0.5),
        axis.text.x = element_text(colour="grey", size=rel(1.5), angle=0, hjust=.5, vjust=.5, face="plain"),
        axis.text.y = element_text(colour="grey", size=rel(1.5), angle=0, hjust=.5, vjust=.5, face="plain"),
        panel.background = element_blank(),
        axis.line = element_line(colour = "grey")##,
        )

p5.3
```

```{r, echo=FALSE}
p6 = ggplot(dat2, aes(group, fill=group))+geom_bar()+
  theme(
        plot.title = element_text(hjust=0.5, size=rel(1.8)),
        axis.title.x = element_text(size=rel(1.8)),
        axis.title.y = element_text(size=rel(1.8), angle=90, vjust=0.5, hjust=0.5),
        axis.text.x = element_text(colour="grey", size=rel(1.5), angle=0, hjust=.5, vjust=.5, face="plain"),
        axis.text.y = element_text(colour="grey", size=rel(1.5), angle=0, hjust=.5, vjust=.5, face="plain"),
        panel.background = element_blank(),
        axis.line = element_line(colour = "grey")##,
        )
p6
```

# Statistical tests

### Standard t-test

Standard t-test comparing:

- Population 1: Offspring with both parents having 2 cotyledons
- Population 2: Offspring with both parents having more than 2 cotyledons

Assumptions:

- Data distributed normal
- Equal variance in the two groups (we can test this with the Levene's test later)

Plotting the densities:

```{r, echo=FALSE}
dat3 = subset(dat2,group %in% c("2x2","(2+)x(2+)"))
dat3$group = factor(dat3$group)

q1 = ggplot(dat3, aes(cotyledons, fill=group))+geom_density(alpha=0.25)+
  theme(
        plot.title = element_text(hjust=0.5, size=rel(1.8)),
        axis.title.x = element_text(size=rel(1.8)),
        axis.title.y = element_text(size=rel(1.8), angle=90, vjust=0.5, hjust=0.5),
        axis.text.x = element_text(colour="grey", size=rel(1.5), angle=0, hjust=.5, vjust=.5, face="plain"),
        axis.text.y = element_text(colour="grey", size=rel(1.5), angle=0, hjust=.5, vjust=.5, face="plain"),
        panel.background = element_blank(),
        axis.line = element_line(colour = "grey")##,
        )
q1
```

Not really normal!

```{r}
summary(dat2$group)
x = dat2$cotyledons[dat2$group == "2x2"]
y = dat2$cotyledons[dat2$group == "(2+)x(2+)"]
```

Variances look similar:
```{r}
var(x)
var(y)
```

But let's do a Levene test:
```{r}
leveneTest(cotyledons ~ group, dat3, center=mean)
```

The p-value is greater than 0.05, so we do not reject the null hypothesis of equal variances.

The Levene test also assumes Normality. So, we can run the alternative Fligner-Killeen test with the same conclusion (equal variances):
```{r}
fligner.test(cotyledons ~ group, dat3)
```

Back to the t-test assumptions:

- Normality: not met
- Equality of variances: met

If we ignore the violation of normality and we run a standard t-test, we get:
```{r}
t.test(x,y)
```

The p-value is less than 0.05, so we reject the null hypothesis of equality of means. This implies that we have evidence that the number of cotyledons is significantly larger in the offspring from "(2+)x(2+)" than from "2x2".

Our justification to do a standard t-test is that if the sample size is moderately large, the two-sample t-test is robust to non-normality due to the central limit theorem.

### Wilcoxon-Mann-Whitney test

Alternative to the t-test when normality is violated. This test does not test equality of means, but equality of distributions, so the conclusion could be deceiving (that is, two different distributions with the same mean will have a rejected null hypothesis using WMW test).

```{r}
wilcox.test(cotyledons ~ group, dat3)
```

The p-value less than 0.05 allows us to also reject the null hypothesis of equality of distributions which provides evidence that the number of cotyledons on the "(2+)x(2+)" group is significantly higher than in the "2x2" group.

### One-way ANOVA

If we want to compare more than two groups, we can do so with a one-way ANOVA test.
We will compare the number of cotyledons on the four groups:

- Population 1: Parent plants
- Population 2: Offspring 2x2
- Population 3: Offspring 2x(2+)
- Population 4: Offspring (2+)x(2+)

Assumptions:

- The observations are obtained independently and randomly from the population defined by the factor levels
- The data of each factor level are normally distributed.
- These normal populations have a common variance

```{r}
res.aov <- aov(cotyledons ~ group, data = dat2)
summary(res.aov)
```

We reject that all groups have the same mean number of cotyledons.

We can check the assumptions:
```{r}
# Homogeneity of variance
plot(res.aov,1)

# Normality
plot(res.aov,2)
```

### Kruskal-Wallis rank sum test

When normality is not met, we can use a Kruskal-Wallis rank sum test, with the same conclusion:
```{r}
kruskal.test(cotyledons ~ group, data = dat2)
```

### Tukey HSD pairwise tests

We can move on to do Tukey multiple pairwise-comparisons (HSD test):
```{r}
TukeyHSD(res.aov)
 ```

### 2020-07-12 shiny-app/webinar-aug20/app.R
started shiny app

@@ -1,0 +1,61 @@
+library(shiny)
+
+## Only run examples in interactive R sessions
+if (interactive()) {
+  
+  ui <- fluidPage(
+    sidebarLayout(
+      sidebarPanel(
+        fileInput("file1", "Choose CSV File",
+                  accept = c(
+                    "text/csv",
+                    "text/comma-separated-values,text/plain",
+                    ".csv")
+        ),
+        tags$hr(),
+        checkboxInput("header", "Header", TRUE)
+      ),
+      # Main panel for displaying outputs ----
+      mainPanel(
+        
+        # Output: Verbatim text for data summary ----
+        verbatimTextOutput("summary"),
+        
+        # Output: HTML table with requested number of observations ----
+        tableOutput("view")
+        
+      )
+    )
+  )
+  
+  server <- function(input, output) {
+    output$summary <- renderPrint({
+      # input$file1 will be NULL initially. After the user selects
+      # and uploads a file, it will be a data frame with 'name',
+      # 'size', 'type', and 'datapath' columns. The 'datapath'
+      # column will contain the local filenames where the data can
+      # be found.
+      inFile <- input$file1
+      
+      if (is.null(inFile))
+        return("No file found")
+      
+      dataset <- read.csv(inFile$datapath, header = input$header)
+      summary(dataset)
+    })
+    
+    # Show the first "n" observations ----
+    output$view <- renderTable({
+      inFile <- input$file1
+      
+      if (is.null(inFile))
+        return(NULL)
+      
+      dataset <- read.csv(inFile$datapath, header = input$header)
+      head(dataset, n = 10)
+    })
+  }
+  
+  shinyApp(ui, server)
+}
+

### 2020-07-12 shiny-app/webinar-aug20/app.R
updated upload

@@ -1,61 +1,124 @@
 library(shiny)
 
-## Only run examples in interactive R sessions
-if (interactive()) {
+# Define UI for data upload app ----
+ui <- fluidPage(
   
-  ui <- fluidPage(
-    sidebarLayout(
-      sidebarPanel(
-        fileInput("file1", "Choose CSV File",
-                  accept = c(
-                    "text/csv",
-                    "text/comma-separated-values,text/plain",
-                    ".csv")
-        ),
-        tags$hr(),
-        checkboxInput("header", "Header", TRUE)
-      ),
-      # Main panel for displaying outputs ----
-      mainPanel(
-        
-        # Output: Verbatim text for data summary ----
-        verbatimTextOutput("summary"),
-        
-        # Output: HTML table with requested number of observations ----
-        tableOutput("view")
-        
-      )
-    )
-  )
+  # App title ----
+  titlePanel("Upload Files"),
   
-  server <- function(input, output) {
-    output$summary <- renderPrint({
-      # input$file1 will be NULL initially. After the user selects
-      # and uploads a file, it will be a data frame with 'name',
-      # 'size', 'type', and 'datapath' columns. The 'datapath'
-      # column will contain the local filenames where the data can
-      # be found.
-      inFile <- input$file1
-      
-      if (is.null(inFile))
-        return("No file found")
-      
-      dataset <- read.csv(inFile$datapath, header = input$header)
-      summary(dataset)
-    })
+  # Sidebar layout with input and output definitions ----
+  sidebarLayout(
     
-    # Show the first "n" observations ----
-    output$view <- renderTable({
-      inFile <- input$file1
+    # Sidebar panel for inputs ----
+    sidebarPanel(
       
-      if (is.null(inFile))
-        return(NULL)
+      # Input: Select a file ----
+      fileInput("file1", "Choose CSV File",
+                multiple = FALSE,
+                accept = c("text/csv",
+                           "text/comma-separated-values,text/plain",
+                           ".csv")),
       
-      dataset <- read.csv(inFile$datapath, header = input$header)
-      head(dataset, n = 10)
-    })
-  }
+      # Horizontal line ----
+      tags$hr(),
+      
+      # Input: Checkbox if file has header ----
+      checkboxInput("header", "Header", TRUE),
+      
+      # Input: Select separator ----
+      radioButtons("sep", "Separator",
+                   choices = c(Comma = ",",
+                               Semicolon = ";",
+                               Tab = "\t"),
[…]

### 2020-07-13 shiny-app/webinar-aug20/app.R
improved shiny app

@@ -1,16 +1,24 @@
 library(shiny)
+# to do:
+# - read the df just once, not in every render function
+# - robust to column names
+# - different tabs? many more options for the user
 
 # Define UI for data upload app ----
 ui <- fluidPage(
   
   # App title ----
-  titlePanel("Upload Files"),
+  #titlePanel("FastStats"),
   
   # Sidebar layout with input and output definitions ----
   sidebarLayout(
     
     # Sidebar panel for inputs ----
     sidebarPanel(
+      
+      img(src = "logo.png", height = 75, width = 230),
+      
+      h3("Data Upload:"),
       
       # Input: Select a file ----
       fileInput("file1", "Choose CSV File",
@@ -35,30 +43,95 @@ ui <- fluidPage(
       # Horizontal line ----
       tags$hr(),
       
+      h3("Data Summary:"),
+      
       # Input: Select number of rows to display ----
       radioButtons("disp", "Display",
                    choices = c(Head = "head",
                                All = "all"),
-                   selected = "head")
+                   selected = "head"),
+      
+      actionButton("goSummary", "Summary"),
+      
+      # Horizontal line ----
+      tags$hr(),
+      
+      h3("Data Visualization:"),
+      
+      actionButton("goHist", "Histogram"),
+      
+      actionButton("goViolin", "Violin Plot"),
+ 
+      # Horizontal line ----
+      tags$hr(),
+      
+      h3("Data Analysis:"),
+      
+      actionButton("goT", "t test"),     
       
     ),
     
     # Main panel for displaying outputs ----
     mainPanel(
       
+      
       # Output: Verbatim text for data summary ----
       verbatimTextOutput("summary"),
       
       # Output: Data file ----
-      tableOutput("contents")
+      tableOutput("contents"),
+      
+      # Output: t test ----
+      verbatimTextOutput("ttest"),
+      
+      # Output: plot
+      plotOutput("histogram"),
+      
+      # Output: plot
+      plotOutput("violinPlot"),
+      
+      
+      
       
     )
     
   )
 )
 
+library(ggplot2)
 # Define server logic to read selected file ----
 server <- function(input, output) {
+  v <- reactiveValues(doViolinPlot = FALSE)
+  
+  observeEvent(input$goViolin, {
[…]

### 2020-07-18 shiny-app/webinar-aug20/server.R
enable the plots work as expected; split app.R to ui.R and server.R; add user-friendly texts

@@ -1,107 +1,39 @@
-library(shiny)
-# to do:
-# - read the df just once, not in every render function
-# - robust to column names
-# - different tabs? many more options for the user
-
-# Define UI for data upload app ----
-ui <- fluidPage(
-  
-  # App title ----
-  #titlePanel("FastStats"),
-  
-  # Sidebar layout with input and output definitions ----
-  sidebarLayout(
-    
-    # Sidebar panel for inputs ----
-    sidebarPanel(
-      
-      img(src = "logo.png", height = 75, width = 230),
-      
-      h3("Data Upload:"),
-      
-      # Input: Select a file ----
-      fileInput("file1", "Choose CSV File",
-                multiple = FALSE,
-                accept = c("text/csv",
-                           "text/comma-separated-values,text/plain",
-                           ".csv")),
-      
-      # Horizontal line ----
-      tags$hr(),
-      
-      # Input: Checkbox if file has header ----
-      checkboxInput("header", "Header", TRUE),
-      
-      # Input: Select separator ----
-      radioButtons("sep", "Separator",
-                   choices = c(Comma = ",",
-                               Semicolon = ";",
-                               Tab = "\t"),
-                   selected = ","),
-      
-      # Horizontal line ----
-      tags$hr(),
-      
-      h3("Data Summary:"),
-      
-      # Input: Select number of rows to display ----
-      radioButtons("disp", "Display",
-                   choices = c(Head = "head",
-                               All = "all"),
-                   selected = "head"),
-      
-      actionButton("goSummary", "Summary"),
-      
-      # Horizontal line ----
-      tags$hr(),
-      
-      h3("Data Visualization:"),
-      
-      actionButton("goHist", "Histogram"),
-      
-      actionButton("goViolin", "Violin Plot"),
- 
-      # Horizontal line ----
-      tags$hr(),
-      
-      h3("Data Analysis:"),
-      
-      actionButton("goT", "t test"),     
-      
-    ),
-    
-    # Main panel for displaying outputs ----
-    mainPanel(
-      
-      
-      # Output: Verbatim text for data summary ----
-      verbatimTextOutput("summary"),
-      
-      # Output: Data file ----
-      tableOutput("contents"),
[…]

### 2020-07-18 shiny-app/webinar-aug20/ui.R

@@ -1,0 +1,107 @@
+library(shiny)
+# to do:
+# - read the df just once, not in every render function
+# - robust to column names
+# - different tabs? many more options for the user
+
+# Define UI for data upload app ----
+ui <- fluidPage(
+  
+  # App title ----
+  #titlePanel("FastStats"),
+  
+  # Sidebar layout with input and output definitions ----
+  sidebarLayout(
+    
+    # Sidebar panel for inputs ----
+    sidebarPanel(
+      
+      img(src = "logo.png", height = 75, width = 230),
+      
+      h3("Data Upload:"),
+      
+      # Input: Select a file ----
+      fileInput("file1", "Choose CSV File or Drag the file here",
+                multiple = FALSE,
+                accept = c("text/csv",
+                           "text/comma-separated-values,text/plain",
+                           ".csv")),
+      
+      # Horizontal line ----
+      tags$hr(),
+      
+      # Input: Checkbox if file has header ----
+      tags$p("Click the checkbox if file has a header:"),
+      checkboxInput("header", "Header", TRUE),
+      
+      # Input: Select separator ----
+      radioButtons("sep", "Choose the separator",
+                   choices = c(Comma = ",",
+                               Semicolon = ";",
+                               Tab = "\t"),
+                   selected = ","),
+     # seems like selected is okay
+      
+      
+      # Horizontal line ----
+      tags$hr(),
+      
+      h3("Data Summary:"),
+      
+      # Input: Select number of rows to display ----
+      radioButtons("disp", "Display the data",
+                   choices = c(Head = "head",
+                               All = "all"),
+                   selected = "head"),
+      # FIXME: users click on the checkbox and the app changes immediately, needs to let it change after click the summary button; can be solved it by isolated()
+      actionButton("goSummary", "Summary"),
+      
+      # Horizontal line ----
+      tags$hr(),
+      
+      h3("Data Visualization:"),
+    
+     fluidRow(
+       column(6,radioButtons("xaxisGrp","X-Axis:", c("1"="1","2"="2"))),
+       column(6,radioButtons("yaxisGrp","Y-axis:", c("1"="1","2"="2")))
+     ),
+      
+      actionButton("goHist", "Histogram"),
+      
+      actionButton("goViolin", "Violin Plot"),
+ 
+      # Horizontal line ----
+      tags$hr(),
+      
+      h3("Data Analysis:"),
+      
[…]

### 2020-07-18 cotyledon-data-analysis.Rmd
updated Rmd with the new data; added some notes about the shiny app

@@
 ```
+
+# Initial data: `Polycot-data-from-NY.csv`
 
@@
 
-# Data visualization
+## Data visualization
 
@@
 
-# Statistical tests
+## Statistical tests
 
@@
 ```
+
+# New data: `Polycot-data-2020.csv`
+
+### Reading the data
+
+Reading in the data (had to remove manually the '#' sign from the column and save as csv file):
+```{r}
+dat = read.table("data/Polycot-data-2020.csv", sep=",",header=TRUE)
+str(dat)
+head(dat)
+summary(dat)
+dat = within(dat, Generation<-factor(Generation, levels=c("P","O")))
+```
+
+### Data visualization
+
+```{r, echo=FALSE}
+p1 = ggplot(dat, aes(Generation, fill=Generation))+geom_bar()+
+  theme(
+        plot.title = element_text(hjust=0.5, size=rel(1.8)),
+        axis.title.x = element_text(size=rel(1.8)),
+        axis.title.y = element_text(size=rel(1.8), angle=90, vjust=0.5, hjust=0.5),
+        axis.text.x = element_text(colour="grey", size=rel(1.5), angle=0, hjust=.5, vjust=.5, face="plain"),
+        axis.text.y = element_text(colour="grey", size=rel(1.5), angle=0, hjust=.5, vjust=.5, face="plain"),
+        panel.background = element_blank(),
+        axis.line = element_line(colour = "grey")##,
+        )
+p1
+```
+
+```{r, echo=FALSE}
+p3 = ggplot(dat, aes(y = Generation, x = Cotyledons, fill = Generation)) +
+  geom_jitter(pch = 21, alpha=0.3, width=0.2)+
+  theme(
+        plot.title = element_text(hjust=0.5, size=rel(1.8)),
+        axis.title.x = element_text(size=rel(1.8)),
+        axis.title.y = element_text(size=rel(1.8), angle=90, vjust=0.5, hjust=0.5),
+        axis.text.x = element_text(colour="grey", size=rel(1.5), angle=0, hjust=.5, vjust=.5, face="plain"),
+        axis.text.y = element_text(colour="grey", size=rel(1.5), angle=0, hjust=.5, vjust=.5, face="plain"),
+        panel.background = element_blank(),
+        axis.line = element_line(colour = "grey")##,
+        )
+p3
+```
+
+```{r, echo=FALSE}
+p3.2 = ggplot(dat, aes(x = Generation, y = Cotyledons, fill = Generation)) +
+  geom_jitter(pch = 21, alpha=0.3, height=0.2)+
+    geom_boxplot(outlier.size = 0, alpha=0.1) +
+  theme(
+        plot.title = element_text(hjust=0.5, size=rel(1.8)),
+        axis.title.x = element_text(size=rel(1.8)),
+        axis.title.y = element_text(size=rel(1.8), angle=90, vjust=0.5, hjust=0.5),
+        axis.text.x = element_text(colour="grey", size=rel(1.5), angle=0, hjust=.5, vjust=.5, face="plain"),
+        axis.text.y = element_text(colour="grey", size=rel(1.5), angle=0, hjust=.5, vjust=.5, face="plain"),
+        panel.background = element_blank(),
+        axis.line = element_line(colour = "grey")##,
+        )
+p3.2
+```
+
+```{r, echo=FALSE}
+p4 = ggplot(dat, aes(x=Generation, y=Cotyledons, fill=Generation))+geom_violin(alpha=0.5)+
+  ylim(c(1,6))+
+  theme(
+        plot.title = element_text(hjust=0.5, size=rel(1.8)),
+        axis.title.x = element_text(size=rel(1.8)),
+        axis.title.y = element_text(size=rel(1.8), angle=90, vjust=0.5, hjust=0.5),
+        axis.text.x = element_text(colour="grey", size=rel(1.5), angle=0, hjust=.5, vjust=.5, face="plain"),
+        axis.text.y = element_text(colour="grey", size=rel(1.5), angle=0, hjust=.5, vjust=.5, face="plain"),
+        panel.background = element_blank(),
+        axis.line = element_line(colour = "grey")##,
+        )
+
+p4
+```
+
+```{r, echo=FALSE}
+p4.2 = ggplot(dat, aes(x=Generation, y=Cotyledons, fill=Generation))+geom_violin(alpha=1)+
+    geom_jitter(pch = 21, alpha=0.1)+
+  ylim(c(1,6))+
+  theme(
+        plot.title = element_text(hjust=0.5, size=rel(1.8)),
+        axis.title.x = element_text(size=rel(1.8)),
+        axis.title.y = element_text(size=rel(1.8), angle=90, vjust=0.5, hjust=0.5),
+        axis.text.x = element_text(colour="grey", size=rel(1.5), angle=0, hjust=.5, vjust=.5, face="plain"),
+        axis.text.y = element_text(colour="grey", size=rel(1.5), angle=0, hjust=.5, vjust=.5, face="plain"),
+        panel.background = element_blank(),
+        axis.line = element_line(colour = "grey")##,
+        )
+
+p4.2
+```
+
+## Statistical tests
+
+### Standard t-test
+
+Standard t-test comparing:
+
+- Population 1: Parents
+- Population 2: Offspring
+
+Assumptions:
+
+- Data distributed normal
+- Equal variance in the two groups (we can test this with the Levene's test later)
+
+Plotting the densities:
+
+```{r, echo=FALSE}
+q1 = ggplot(dat, aes(Cotyledons, fill=Generation))+geom_density(alpha=0.25)+
+  theme(
+        plot.title = element_text(hjust=0.5, size=rel(1.8)),
+        axis.title.x = element_text(size=rel(1.8)),
+        axis.title.y = element_text(size=rel(1.8), angle=90, vjust=0.5, hjust=0.5),
+        axis.text.x = element_text(colour="grey", size=rel(1.5), angle=0, hjust=.5, vjust=.5, face="plain"),
+        axis.text.y = element_text(colour="grey", size=rel(1.5), angle=0, hjust=.5, vjust=.5, face="plain"),
+        panel.background = element_blank(),
+        axis.line = element_line(colour = "grey")##,
+        )
+q1
+```
+
+Not really normal!
+
+```{r}
+x = dat$Cotyledons[dat$Generation == "P"]
+y = dat$Cotyledons[dat$Generation == "O"]
+```
+
+Variances look similar:
+```{r}
+var(x)
+var(y)
+```
+
+But let's do a Levene test:
+```{r}
+leveneTest(Cotyledons ~ Generation, dat, center=mean)
+```
+
+The p-value is less than 0.05, so we reject the null hypothesis of equal variances.
+
+The Levene test also assumes Normality. So, we can run the alternative Fligner-Killeen test with the same conclusion (equal variances):
+```{r}
+fligner.test(Cotyledons ~ Generation, dat)
+```
+
+Back to the t-test assumptions:
+
+- Normality: not met
+- Equality of variances: not met
+
+If we ignore the violations and we run a standard t-test, we get:
+```{r}
+t.test(x,y, var.equal=TRUE)
+```
+
+The p-value is less than 0.05, so we reject the null hypothesis of equality of means.
+This implies that we have evidence that the number of cotyledons is significantly larger in the offspring than in the parents.
+
+Our justification to do a standard t-test is that if the sample size is moderately large, the two-sample t-test is robust to non-normality due to the central limit theorem.
+
+We can also run the Welch version of t test that does not assume equal variances (this is the default in R):
+
+```{r}
+t.test(x,y)
+```
+
+### Wilcoxon-Mann-Whitney test
+
+Alternative to the t-test when normality is violated.
+This test does not test equality of means, but equality of distributions, so the conclusion could be deceiving (that is, two different distributions with the same mean will have a rejected null hypothesis using WMW test).
+
+```{r}
+wilcox.test(Cotyledons ~ Generation, dat)
+```
+
+The p-value less than 0.05 allows us to also reject the null hypothesis of equality of distributions which provides evidence that the number of cotyledons on the offspring group is significantly higher than in the parent group.

### 2020-07-19 shiny-app/webinar-aug20/server.R
update T-Test part: enable user-input data; but the input format still need to get decided

@@ -14,9 +14,9 @@ server <- function(input, output,session) {
   })
   
 
-  output$contents <- renderTable({
-    data_set()
-  })
+  # output$contents <- renderTable({
+  #   data_set()
+  # })
   observe({
     dsnames <- names(data_set())
     cb_options <- list()
@@ -27,6 +27,14 @@ server <- function(input, output,session) {
                        selected = "")
     updateRadioButtons(session, "yaxisGrp",
                              label = "Y-Axis",
+                             choices = cb_options,
+                             selected = "")
+    updateRadioButtons(session, "groupVar",
+                       label = "Group Variable",
+                       choices = cb_options,
+                       selected = "")
+    updateRadioButtons(session, "quantity",
+                             label = "Quantity",
                              choices = cb_options,
                              selected = "")
   })
@@ -77,6 +85,14 @@ server <- function(input, output,session) {
   v6<-reactiveValues(yv=0)
   observeEvent(input$yaxisGrp,{
     v6$yv<-input$yaxisGrp
+  })
+  v7<-reactiveValues(gv=0)
+  observeEvent(input$groupVar,{
+    v7$gv<-input$groupVar
+  })
+  v8<-reactiveValues(q=0)
+  observeEvent(input$quantity,{
+    v8$q<-input$quantity
   })
   
   output$summary <- renderPrint({
@@ -130,18 +146,31 @@ server <- function(input, output,session) {
       }
     )
     
+    if (v4$doT == FALSE) return()
+    
+    # extract var as col obj
+    # print(v7$gv)
+   
+    group_val<-unlist(subset(df, select=c(v7$gv)))
+    
+    quantity<-unlist(subset(df, select=c(v8$q)))
+
     dat2 = within(df, group <- "(2+)x(2+)")
-    dat2$group[dat2$parents == ""] = "parent"
-    dat2$group[dat2$parents == "2x2"] = "2x2"
-    dat2$group[dat2$parents == "2x3"] = "2x(2+)"
-    dat2$group[dat2$parents == "2x4"] = "2x(2+)"
+    # #############################################
+    dat2$group[subset(dat2, select=c(v7$gv)) == ""] = "parent"
[…]

### 2020-07-19 shiny-app/webinar-aug20/ui.R

@@ -74,6 +74,11 @@ ui <- fluidPage(
       tags$hr(),
       
       h3("Data Analysis:"),
+     
+     fluidRow(
+       column(6,radioButtons("groupVar","Group Variable:", c("1"="1","2"="2"))),
+       column(6,radioButtons("quantity","Quantity:", c("1"="1","2"="2")))
+     ),
       
       actionButton("goT", "t test"),     
       

### 2020-07-19 shiny-app/webinar-aug20/server.R
Updates on boxplot, scatterplot, densities; complete basic funcs

@@ -22,11 +22,11 @@ server <- function(input, output,session) {
     cb_options <- list()
     cb_options[ dsnames] <- dsnames
     updateRadioButtons(session, "xaxisGrp",
-                       label = "X-Axis",
+                       label = "Grouping Variables",
                        choices = cb_options,
                        selected = "")
     updateRadioButtons(session, "yaxisGrp",
-                             label = "Y-Axis",
+                             label = "Quantity",
                              choices = cb_options,
                              selected = "")
     updateRadioButtons(session, "groupVar",
@@ -52,7 +52,6 @@ server <- function(input, output,session) {
   v2 <- reactiveValues(doHist = FALSE)
   # serve as a middleman between input and output, we can see that render func use v2$doHist instead of input$goHist
   # I guess the id val (i'm not sure about id's class) can only be assigned to bool val; surprisingly, it's int val; so that means R can also treat int as bool used in conditional expr (like C; FALSE equals to 0)
-  
   observeEvent(input$goHist, {
     # 0 will be coerced to FALSE
     # 1+ will be coerced to TRUE
@@ -62,7 +61,6 @@ server <- function(input, output,session) {
   })
   
   v3 <- reactiveValues(doSummary = FALSE)
-  
   observeEvent(input$goSummary, {
     # 0 will be coerced to FALSE
     # 1+ will be coerced to TRUE
@@ -70,8 +68,6 @@ server <- function(input, output,session) {
   })
   
   v4 <- reactiveValues(doT = FALSE)
-  
-  
   observeEvent(input$goT, {
     # 0 will be coerced to FALSE
     # 1+ will be coerced to TRUE
@@ -82,17 +78,41 @@ server <- function(input, output,session) {
   observeEvent(input$xaxisGrp,{
     v5$xv<-input$xaxisGrp
   })
+  
   v6<-reactiveValues(yv=0)
   observeEvent(input$yaxisGrp,{
     v6$yv<-input$yaxisGrp
   })
+  
   v7<-reactiveValues(gv=0)
   observeEvent(input$groupVar,{
     v7$gv<-input$groupVar
   })
+  
   v8<-reactiveValues(q=0)
   observeEvent(input$quantity,{
     v8$q<-input$quantity
+  })
+  
+  v9 <- reactiveValues(doScatter = FALSE)
+  observeEvent(input$goScatter, {
+    # 0 will be coerced to FALSE
+    # 1+ will be coerced to TRUE
[…]

### 2020-07-19 shiny-app/webinar-aug20/ui.R

@@ -49,7 +49,7 @@ ui <- fluidPage(
       h3("Data Summary:"),
       
       # Input: Select number of rows to display ----
-      radioButtons("disp", "Display the data",
+      radioButtons("disp", "Data Customization",
                    choices = c(Head = "head",
                                All = "all"),
                    selected = "head"),
@@ -62,13 +62,19 @@ ui <- fluidPage(
       h3("Data Visualization:"),
     
      fluidRow(
-       column(6,radioButtons("xaxisGrp","X-Axis:", c("1"="1","2"="2"))),
-       column(6,radioButtons("yaxisGrp","Y-axis:", c("1"="1","2"="2")))
+       column(6,radioButtons("xaxisGrp","Grouping Variables:", c("1"="1","2"="2"))),
+       column(6,radioButtons("yaxisGrp","Quantity:", c("1"="1","2"="2")))
      ),
       
       actionButton("goHist", "Histogram"),
       
       actionButton("goViolin", "Violin Plot"),
+     
+      actionButton("goScatter", "Scatter Plot"),
+     
+      actionButton("goBox", "Box Plot"),
+     
+      actionButton("goDensities", "Densities"),
  
       # Horizontal line ----
       tags$hr(),
@@ -102,6 +108,12 @@ ui <- fluidPage(
       
       # Output: plot
       plotOutput("violinPlot"),
+      
+      plotOutput("scatterPlot"),
+      
+      plotOutput("boxPlot"),
+      
+      plotOutput("densities"),
       
       
       

### 2020-07-19 shiny-app/webinar-aug20/server.R
enable to plot with/without dots; enable to adjust plot size

@@ -5,7 +5,7 @@ server <- function(input, output,session) {
   v <- reactiveValues(doViolinPlot = FALSE)
   # #################################### 
   dsnames<-c() #a vector to store col names
-    
+  
   data_set <- reactive({
     req(input$file1)
     inFile <- input$file1
@@ -13,7 +13,7 @@ server <- function(input, output,session) {
                        sep=input$sep)
   })
   
-
+  
   # output$contents <- renderTable({
   #   data_set()
   # })
@@ -22,21 +22,21 @@ server <- function(input, output,session) {
     cb_options <- list()
     cb_options[ dsnames] <- dsnames
     updateRadioButtons(session, "xaxisGrp",
-                       label = "Grouping Variables",
+                       label = "Group Variable",
                        choices = cb_options,
                        selected = "")
     updateRadioButtons(session, "yaxisGrp",
-                             label = "Quantity",
-                             choices = cb_options,
-                             selected = "")
+                       label = "Quantity",
+                       choices = cb_options,
+                       selected = "")
     updateRadioButtons(session, "groupVar",
                        label = "Group Variable",
                        choices = cb_options,
                        selected = "")
     updateRadioButtons(session, "quantity",
-                             label = "Quantity",
-                             choices = cb_options,
-                             selected = "")
+                       label = "Quantity",
+                       choices = cb_options,
+                       selected = "")
   })
   output$choose_dataset <- renderUI({
     selectInput("dataset", "Data set", as.list(data_sets))
@@ -56,8 +56,6 @@ server <- function(input, output,session) {
     # 0 will be coerced to FALSE
     # 1+ will be coerced to TRUE
     v2$doHist <- input$goHist
-    print(class(v2$doHist))
-    
   })
   
   v3 <- reactiveValues(doSummary = FALSE)
@@ -115,6 +113,35 @@ server <- function(input, output,session) {
     v11$doDensities <- input$goDensities
   })
   
[…]

### 2020-07-19 shiny-app/webinar-aug20/ui.R

@@ -40,7 +40,7 @@ ui <- fluidPage(
                                Semicolon = ";",
                                Tab = "\t"),
                    selected = ","),
-     # seems like selected is okay
+      # seems like selected is okay
       
       
       # Horizontal line ----
@@ -59,32 +59,50 @@ ui <- fluidPage(
       # Horizontal line ----
       tags$hr(),
       
+      h3("Plot Option:"),
+      
+      checkboxInput("addPoints", "Add data points", T),
+      
+      checkboxInput("adjustSize", "Adjust plot size", F),
+      
+      conditionalPanel(
+        # only show this panel when user wanna adjust plot size; this panel uses JS, don't need to have a quotation mark over true
+        condition="input.adjustSize==true",
+        numericInput("widthVal","Plot Width:",550),
+        numericInput("heightVal","Plot Height:",400)
+        # I've tested the result and seems that (550,400) is a great parameter for this, if val gets too large, plot simply go out of border in the mainpanel
+        # numericInput is the specific input branket for integer inputs; don't use textinput in this case, cos id$val will be characters instead of string
+      ),
+      # options func may solve this 
+      
+      tags$hr(),
+      
       h3("Data Visualization:"),
-    
-     fluidRow(
-       column(6,radioButtons("xaxisGrp","Grouping Variables:", c("1"="1","2"="2"))),
-       column(6,radioButtons("yaxisGrp","Quantity:", c("1"="1","2"="2")))
-     ),
+      
+      fluidRow(
+        column(6,radioButtons("xaxisGrp","Grouping Variables:", c("1"="1","2"="2"))),
+        column(6,radioButtons("yaxisGrp","Quantity:", c("1"="1","2"="2")))
+      ),
       
       actionButton("goHist", "Histogram"),
       
       actionButton("goViolin", "Violin Plot"),
-     
+      
       actionButton("goScatter", "Scatter Plot"),
-     
+      
       actionButton("goBox", "Box Plot"),
-     
+      
       actionButton("goDensities", "Densities"),
- 
+      
       # Horizontal line ----
       tags$hr(),
       
       h3("Data Analysis:"),
-     
-     fluidRow(
-       column(6,radioButtons("groupVar","Group Variable:", c("1"="1","2"="2"))),
-       column(6,radioButtons("quantity","Quantity:", c("1"="1","2"="2")))
-     ),
+      
+      fluidRow(
[…]

### 2020-07-20 shiny-app/webinar-aug20/server.R
enable all the tests work as expected

@@ -1,5 +1,6 @@
 library(data.table)
 library(ggplot2)
+library(car)
 # Define server logic to read selected file ----
 server <- function(input, output,session) {
   v <- reactiveValues(doViolinPlot = FALSE)
@@ -127,21 +128,48 @@ server <- function(input, output,session) {
     v13$doAdjustSize<- input$adjustSize
   })
   
-  # v14 <- reactiveValues(doWidthVal =700)
-  # observeEvent(input$widthVal, {
-  #   # 0 will be coerced to FALSE
-  #   # 1+ will be coerced to TRUE
-  #   v14$doWidthVal<- input$widthVal
-  # })
-  # 
-  # v15 <- reactiveValues(doHeightVal =550)
-  # observeEvent(input$heightVal, {
-  #   # 0 will be coerced to FALSE
-  #   # 1+ will be coerced to TRUE
-  #   v15$doHeightVal<- input$heightVal
-  # })
+  v14 <- reactiveValues(sel1 =0)
+  observeEvent(input$group1, {
+    # 0 will be coerced to FALSE
+    # 1+ will be coerced to TRUE
+    v14$sel1<- input$group1
+  })
   
+  v15 <- reactiveValues(sel2 =0)
+  observeEvent(input$group2, {
+    # 0 will be coerced to FALSE
+    # 1+ will be coerced to TRUE
+    v15$sel2<- input$group2
+  })
   
+  v16 <- reactiveValues(doEqualVar=F)
+  observeEvent(input$isEqualVar, {
+    # 0 will be coerced to FALSE
+    # 1+ will be coerced to TRUE
+    v16$doEqualVar<- input$isEqualVar
+  })
+  
+  v17 <- reactiveValues(doLevene = FALSE)
+  observeEvent(input$goLevene, {
+    # 0 will be coerced to FALSE
+    # 1+ will be coerced to TRUE
+    v17$doLevene <- input$goLevene
+  })
+  
+  v18 <- reactiveValues(doFligner = FALSE)
+  observeEvent(input$goFligner, {
+    # 0 will be coerced to FALSE
+    # 1+ will be coerced to TRUE
+    v18$doFligner <- input$goFligner
+  })
+  
+  v19 <- reactiveValues(doWilcoxon = FALSE)
+  observeEvent(input$goWilcoxon, {
+    # 0 will be coerced to FALSE
+    # 1+ will be coerced to TRUE
+    v19$doWilcoxon <- input$goWilcoxon
+  })
+   
[…]

### 2020-07-20 shiny-app/webinar-aug20/ui.R

@@ -97,14 +97,30 @@ ui <- fluidPage(
       # Horizontal line ----
       tags$hr(),
       
+      h3("Data Analysis Option:"),
+      
+      checkboxInput("isEqualVar", "Equal Variance", F),
+      
+      tags$hr(),
+      
       h3("Data Analysis:"),
       
       fluidRow(
-        column(6,radioButtons("groupVar","Group Variable:", c("1"="1","2"="2"))),
-        column(6,radioButtons("quantity","Quantity:", c("1"="1","2"="2")))
+        column(6,radioButtons("groupVar","Group Variable:", c("1"="UnSpecified_Value","2"="UnSpecified_Value"))),
+        column(6,radioButtons("quantity","Quantity:", c("1"="UnSpecified_Value","2"="UnSpecified_Value")))
       ),
       
-      actionButton("goT", "t test"),     
+      uiOutput("sel1"),
+      
+      uiOutput("sel2"),
+      
+      actionButton("goT", "T test"), 
+      
+      actionButton("goLevene", "Levene test"),
+      
+      actionButton("goFligner", "Fligner test"),  
+      
+      actionButton("goWilcoxon", "Wilcoxon test"),     
       
     ),
     
@@ -120,6 +136,12 @@ ui <- fluidPage(
       
       # Output: t test ----
       verbatimTextOutput("ttest"),
+      
+      verbatimTextOutput("levene"),
+      
+      verbatimTextOutput("fligner"),
+      
+      verbatimTextOutput("wilcoxon"),
       
       # Output: plot
       plotOutput("histogram"),

### 2020-07-20 shiny-app/webinar-aug20/server.R
add error suppress tag

@@ -10,8 +10,9 @@ server <- function(input, output,session) {
   data_set <- reactive({
     req(input$file1)
     inFile <- input$file1
-    data_set<-read.csv(inFile$datapath, header=input$header, 
-                       sep=input$sep)
+    read.csv(inFile$datapath, header=input$header, 
+                         sep=input$sep)
+   
   })
   
   
@@ -316,8 +317,14 @@ server <- function(input, output,session) {
     if (v2$doHist == FALSE) return()
     
     # print(subset(df, select=c(v5$xv)))
-    
+  
+    # tryCatch({
+    #   x_val<-unlist(subset(df, select=c(v5$xv)))
+    # }error=function(e){
+    #   stop(safeError(e))
+    # })
     x_val<-unlist(subset(df, select=c(v5$xv)))
+    
     # notice that v5$xv here is character, not col obj; but ggplot needs to accept col obj
     # use subset func to extract col object from dataframe; other func, such as df[...], seems not work at all
     # ggplot2 does not accept list object; must use unlist
@@ -337,8 +344,7 @@ server <- function(input, output,session) {
         axis.line = element_line(colour = "grey")##,
       )    
     
-    
-  },,width=exprToFunction(input$widthVal),height=exprToFunction(input$heightVal))
+  },width=exprToFunction(input$widthVal),height=exprToFunction(input$heightVal))
   
   output$scatterPlot<-renderPlot({
     # FIXME:we want to read the input file only once per session
@@ -487,7 +493,7 @@ server <- function(input, output,session) {
       )
     
     
-  },,width=exprToFunction(input$widthVal),height=exprToFunction(input$heightVal))
+  },width=exprToFunction(input$widthVal),height=exprToFunction(input$heightVal))
   
   output$sel1<-renderUI({
     req(input$file1)

### 2020-07-20 shiny-app/webinar-aug20/ui.R

@@ -155,8 +155,11 @@ ui <- fluidPage(
       
       plotOutput("densities"),
       
-      
-      
+      # used to suppress the errors; comment it out everytime we need to debug
+      tags$style(type="text/css",
+                 ".shiny-output-error { visibility: hidden; }",
+                 ".shiny-output-error:before { visibility: hidden; }"
+      )
       
     )
     

### 2020-07-21 cotyledon-data-analysis.Rmd
minor change to plot

@@
 ```{r, echo=FALSE}
-p3 = ggplot(dat, aes(y = Generation, x = Cotyledons, fill = Generation)) +
-  geom_jitter(pch = 21, alpha=0.3, width=0.2)+
+p3 = ggplot(dat, aes(x = Generation, y = Cotyledons, fill = Generation)) +
+  geom_jitter(pch = 21, alpha=0.3, height=0.2)+
   theme(

### 2020-07-23 shiny-app/webinar-aug20/server.R
use Plotly to produce all the plots; solve zoom in zoom out problem; able to download with png format;enable far more interactive(dynamic) plots than common plot methods

@@ -1,6 +1,7 @@
 library(data.table)
 library(ggplot2)
 library(car)
+library(plotly)
 # Define server logic to read selected file ----
 server <- function(input, output,session) {
   v <- reactiveValues(doViolinPlot = FALSE)
@@ -231,7 +232,7 @@ server <- function(input, output,session) {
     
   })
   
-  output$violinPlot <- renderPlot({
+  output$violinPlot <- renderPlotly({
     
     # there must be a way not to repeat the same lines to get df
     
@@ -280,17 +281,18 @@ server <- function(input, output,session) {
         axis.text.x = element_text(colour="grey", size=rel(1.5), angle=0, hjust=.5, vjust=.5, face="plain"),
         axis.text.y = element_text(colour="grey", size=rel(1.5), angle=0, hjust=.5, vjust=.5, face="plain"),
         panel.background = element_blank(),
+        text=element_text(size=8),
         axis.line = element_line(colour = "grey")##,
       )  
     if(v12$doAddPoints==T){
-      plot<-plot+geom_point(pch = 21, alpha=0.3, position = position_jitterdodge(jitter.height=0.05, jitter.width=2.5))
+      plot<-plot+geom_point(pch = 21, alpha=0.3, position = position_jitterdodge(jitter.height=0.05, jitter.width=2.5),size=0.5)
     }
-    return (plot)
-  },,width=exprToFunction(input$widthVal),height=exprToFunction(input$heightVal))
+    ggplotly(plot)
+  })
   
   # histogram should be the base template
   # all other plots should follow its styles
-  output$histogram <- renderPlot({
+  output$histogram <- renderPlotly({
     
     # there must be a way not to repeat the same lines to get df
     
@@ -332,7 +334,7 @@ server <- function(input, output,session) {
     # y_val<-df[,v6$yv]
     
     
-    ggplot(df, aes(x_val, fill=x_val))+geom_bar(alpha=0.5)+ 
+    p<-ggplot(df, aes(x_val, fill=x_val))+geom_bar(alpha=0.5)+ 
       xlab(v5$xv)+labs(fill=v5$xv)+
       theme(
         plot.title = element_text(hjust=0.5, size=rel(1.8)),
@@ -341,12 +343,14 @@ server <- function(input, output,session) {
         axis.text.x = element_text(colour="grey", size=rel(1.5), angle=0, hjust=.5, vjust=.5, face="plain"),
[…]

### 2020-07-23 shiny-app/webinar-aug20/ui.R

@@ -1,4 +1,5 @@
 library(shiny)
+library(plotly)
 # to do:
 # - read the df just once, not in every render function
 # - robust to column names
@@ -63,19 +64,22 @@ ui <- fluidPage(
       
       checkboxInput("addPoints", "Add data points", T),
       
-      checkboxInput("adjustSize", "Adjust plot size", F),
+      # ###################################################
+      #if we insist on using plotly, the adjust and controlpanel button is not needed in this appp
+      # checkboxInput("adjustSize", "Adjust plot size", F),
       
-      conditionalPanel(
-        # only show this panel when user wanna adjust plot size; this panel uses JS, don't need to have a quotation mark over true
-        condition="input.adjustSize==true",
-        numericInput("widthVal","Plot Width:",550),
-        numericInput("heightVal","Plot Height:",400)
-        # I've tested the result and seems that (550,400) is a great parameter for this, if val gets too large, plot simply go out of border in the mainpanel
-        # numericInput is the specific input branket for integer inputs; don't use textinput in this case, cos id$val will be characters instead of string
-      ),
+      # conditionalPanel(
+      #   # only show this panel when user wanna adjust plot size; this panel uses JS, don't need to have a quotation mark over true
+      #   condition="input.adjustSize==true",
+      #   numericInput("widthVal","Plot Width:",600),
+      #   numericInput("heightVal","Plot Height:",480)
+      #   # I've tested the result and seems that (550,400) is a great parameter for this, if val gets too large, plot simply go out of border in the mainpanel
+      #   # numericInput is the specific input branket for integer inputs; don't use textinput in this case, cos id$val will be characters instead of string
+      # ),
       # options func may solve this 
       
-      tags$hr(),
+      # tags$hr(),
+      ########################################################
       
       h3("Data Visualization:"),
       
@@ -93,6 +97,10 @@ ui <- fluidPage(
       actionButton("goBox", "Box Plot"),
       
       actionButton("goDensities", "Densities"),
+      
+      tags$hr(),
+      
+      # radioButtons("downloadOptions", "Select the Option", list("png","jpeg","pdf")),
       
       # Horizontal line ----
       tags$hr(),
@@ -121,7 +129,7 @@ ui <- fluidPage(
       actionButton("goFligner", "Fligner test"),  
       
       actionButton("goWilcoxon", "Wilcoxon test"),     
[…]

### 2020-07-24 shiny-app/webinar-aug20/server.R
fix scatterplot bug

@@ -381,7 +381,7 @@ server <- function(input, output,session) {
     y_val<-unlist(subset(df, select=c(v6$yv)))
     
     
-    p<-ggplot(df, aes(y = y_val, x = x_val, fill = y_val)) +
+    p<-ggplot(df, aes(y = y_val, x = x_val, fill = x_val)) +
       xlab(v5$xv)+labs(fill=v5$xv)+ylab(v6$yv)+
       geom_jitter(pch = 21, alpha=0.3, width=0.2)+
       theme(

### 2020-07-25 shiny-app/webinar-aug20/server.R
Uses sidebar UI to build the app; improve some minor elements

@@ -12,8 +12,8 @@ server <- function(input, output,session) {
     req(input$file1)
     inFile <- input$file1
     read.csv(inFile$datapath, header=input$header, 
-                         sep=input$sep)
-   
+             sep=input$sep)
+    
   })
   
   
@@ -171,7 +171,7 @@ server <- function(input, output,session) {
     # 1+ will be coerced to TRUE
     v19$doWilcoxon <- input$goWilcoxon
   })
-   
+  
   output$summary <- renderPrint({
     
     # input$file1 will be NULL initially. After the user selects
@@ -319,7 +319,7 @@ server <- function(input, output,session) {
     if (v2$doHist == FALSE) return()
     
     # print(subset(df, select=c(v5$xv)))
-  
+    
     # tryCatch({
     #   x_val<-unlist(subset(df, select=c(v5$xv)))
     # }error=function(e){
@@ -381,7 +381,7 @@ server <- function(input, output,session) {
     y_val<-unlist(subset(df, select=c(v6$yv)))
     
     
-    p<-ggplot(df, aes(y = y_val, x = x_val, fill = y_val)) +
+    p<-ggplot(df, aes(y = y_val, x = x_val, fill = x_val)) +
       xlab(v5$xv)+labs(fill=v5$xv)+ylab(v6$yv)+
       geom_jitter(pch = 21, alpha=0.3, width=0.2)+
       theme(
@@ -737,7 +737,7 @@ server <- function(input, output,session) {
   #     
   #   } 
   # )
- 
+  
   
   # if users wanna change the parameter (say change dots), we can first set a var as p<-ggplot(...). Then use if statement to test user's response, add it piece by piece, and finally return p
 }

### 2020-07-25 shiny-app/webinar-aug20/ui.R

@@ -1,177 +1,177 @@
 library(shiny)
 library(plotly)
-# to do:
-# - read the df just once, not in every render function
-# - robust to column names
-# - different tabs? many more options for the user
 
-# Define UI for data upload app ----
-ui <- fluidPage(
+shinyUI(pageWithSidebar(
   
-  # App title ----
-  #titlePanel("FastStats"),
+  headerPanel(
+    img(src = "logo.png", height = 50, width = 154),
+    # fix the adjust ratio 75/230
+    # "Fast-Stats Web Tool",
+    #           tags$head(tags$style(type="text/css", "label.radio { display: inline-block; }", ".radio input[type=\"radio\"] { float: none; }"),
+    #                     tags$style(type="text/css", "select { max-width: 200px; }"),
+    #                     tags$style(type="text/css", "textarea { max-width: 185px; }"),
+    #                     tags$style(type="text/css", ".jslider { max-width: 200px; }"),
+    #                     tags$style(type='text/css', ".well { max-width: 330px; }"),
+    #                     tags$style(type='text/css', ".span4 { max-width: 330px; }")) 
+  ),
   
-  # Sidebar layout with input and output definitions ----
-  sidebarLayout(
-    
-    # Sidebar panel for inputs ----
-    sidebarPanel(
-      
-      img(src = "logo.png", height = 75, width = 230),
-      
-      h3("Data Upload:"),
-      
-      # Input: Select a file ----
-      fileInput("file1", "Choose CSV File or Drag the file here",
-                multiple = FALSE,
-                accept = c("text/csv",
-                           "text/comma-separated-values,text/plain",
-                           ".csv")),
-      
-      # Horizontal line ----
-      tags$hr(),
-      
-      # Input: Checkbox if file has header ----
-      tags$p("Click the checkbox if file has a header:"),
-      checkboxInput("header", "Header", TRUE),
-      
-      # Input: Select separator ----
-      radioButtons("sep", "Choose the separator",
-                   choices = c(Comma = ",",
-                               Semicolon = ";",
-                               Tab = "\t"),
-                   selected = ","),
-      # seems like selected is okay
-      
-      
-      # Horizontal line ----
-      tags$hr(),
-      
-      h3("Data Summary:"),
-      
-      # Input: Select number of rows to display ----
-      radioButtons("disp", "Data Customization",
-                   choices = c(Head = "head",
-                               All = "all"),
-                   selected = "head"),
[…]

### 2020-07-26 cotyledon-data-analysis.Rmd
changes to Rmd

@@
 The p-value less than 0.05 allows us to also reject the null hypothesis of equality of distributions which provides evidence that the number of cotyledons on the offspring group is significantly higher than in the parent group.
+
+# New data: `Polycot-data-2020.csv` focusing on chi-square test
+
+### Reading the data
+
+Reading in the data (had to remove manually the '#' sign from the column and save as csv file):
+```{r}
+dat = read.table("data/Polycot-data-2020.csv", sep=",",header=TRUE)
+str(dat)
+head(dat)
+summary(dat)
+dat = within(dat, Generation<-factor(Generation, levels=c("P","O")))
+```
+
+Creating categorical variable: dycots/polycots
+
+```{r}
+dat2 = data.frame(generation=dat$Generation)
+dat2$cots = "dycot"
+dat2$cots[dat$Cotyledons > 2] = "polycot"
+dat2 = within(dat2, cots<-factor(cots))
+summary(dat2)
+write.table(dat2,file="data/polycot-data-2020-categories.csv", row.names = FALSE, sep=",")
+
+dt = table(dat2$cots, dat2$generation)
+```
+
+### Data visualization
+```{r}
+library(graphics)
+mosaicplot(dt, shade = TRUE, las=2, main="Cotelydons")
+```
+
+### Data analysis
+```{r}
+dt
+ct = chisq.test(dt)
+
+obs = ct$observed
+exp = ct$expected
+
+df = rbind(obs[1,],exp[1,],obs[2,],exp[2,],colSums(obs))
+rownames(df) = c("dycot observed", "dycot expected", "polycot observed", "polycot expected", "total")
+
+r = rowSums(obs)
+ss = sum(r)
+totals = c(r[1],NA,r[2],NA,ss)
+cbind(df,totals)
+ct$statistic
+ct$p.value
+
+## expected values
+2437*2556/5585
+2437*3029/5585
+3148*2556/5585
+3148*3029/5585
+```

### 2020-07-27 shiny-app/webinar-aug20/server.R
Add descriptions

@@ -103,7 +103,7 @@ server <- function(input, output,session) {
   })
   
   v10 <- reactiveValues(doBox = FALSE)
-  observeEvent(input$goBox, {
+  observeEvent(input$goBofx, {
     # 0 will be coerced to FALSE
     # 1+ will be coerced to TRUE
     v10$doBox <- input$goBox

### 2020-07-27 shiny-app/webinar-aug20/ui.R

@@ -4,7 +4,7 @@ library(plotly)
 shinyUI(pageWithSidebar(
   
   headerPanel(
-    img(src = "logo.png", height = 50, width = 154),
+    img(src = "logo.png", height = 75, width = 350),
     # fix the adjust ratio 75/230
     # "Fast-Stats Web Tool",
     #           tags$head(tags$style(type="text/css", "label.radio { display: inline-block; }", ".radio input[type=\"radio\"] { float: none; }"),
@@ -42,14 +42,12 @@ shinyUI(pageWithSidebar(
                                               Tab = "\t"),
                                   selected = ","),
                      
-                     h4("Data Summary:"),
-                     
                      # Input: Select number of rows to display ----
                      radioButtons("disp","Choose to show 'head' or 'all'",
                                   choices = c(Head = "head",
                                               All = "all"),
                                   selected = "head"),
-                     actionButton("goSummary", "Summary"),
+                     # actionButton("goSummary", "Summary"),
                      
                      HTML('<p>Data in <a href="http://en.wikipedia.org/wiki/Delimiter-separated_values">delimited text files </a> can be separated by comma, tab or semicolon. 
 				For example, Excel data can be exported in .csv (comma separated) or .tab (tab separated) format. </p>')
@@ -61,6 +59,8 @@ shinyUI(pageWithSidebar(
                      h4("Plot Option:"),
                      
                      checkboxInput("addPoints", "Add data points", T),
+                     
+                     helpText("This option allows the user to add a scatterplot of the data where each dot corresponds to one observation (row) in the dataset."),
                      
                      h4("Data Visualization:"),
                      
@@ -76,7 +76,7 @@ shinyUI(pageWithSidebar(
                                                tags$style(type='text/css', "#button { vertical-align: middle; height: 50px; width: 100%; font-size: 30px;}")
                                         ),
                                       ),
-                                      helpText("help text")
+                                      helpText("Plot a numerical variable 'Quantity' by groups 'Group variable'. It is similar to the box plot but it also shows the distribution and spread of the data.")
                      ),
                      conditionalPanel(condition="input.plotType==1",
                                       fluidRow(
@@ -85,7 +85,7 @@ shinyUI(pageWithSidebar(
                                                tags$style(type='text/css', "#button { vertical-align: middle; height: 50px; width: 100%; font-size: 30px;}")
                                         )
[…]

### 2020-07-27 shiny-app/webinar-aug20/server.R
add chi-sq test; solve error cases; used notification here;fix some minor bugs

@@ -2,9 +2,9 @@ library(data.table)
 library(ggplot2)
 library(car)
 library(plotly)
+library(pracma)
 # Define server logic to read selected file ----
 server <- function(input, output,session) {
-  v <- reactiveValues(doViolinPlot = FALSE)
   # #################################### 
   dsnames<-c() #a vector to store col names
   
@@ -40,12 +40,28 @@ server <- function(input, output,session) {
                        label = "Quantity",
                        choices = cb_options,
                        selected = "")
+    updateRadioButtons(session, "gv1",
+                       label = "Group Variable 1",
+                       choices = cb_options,
+                       selected = "")
+    updateRadioButtons(session, "gv2",
+                       label = "Group Variable 2",
+                       choices = cb_options,
+                       selected = "")
+    updateRadioButtons(session, "gvBox",
+                       label = "Grouping Variable",
+                       choices = cb_options,
+                       selected = "")
+    updateRadioButtons(session, "qBox",
+                       label = "Quantity",
+                       choices = cb_options,
+                       selected = "")
   })
   output$choose_dataset <- renderUI({
     selectInput("dataset", "Data set", as.list(data_sets))
   })
   ####################################
-  
+  v<-reactiveValues(doViolinPlot=FALSE)
   observeEvent(input$goViolin, {
     # 0 will be coerced to FALSE
     # 1+ will be coerced to TRUE
@@ -103,7 +119,7 @@ server <- function(input, output,session) {
   })
   
   v10 <- reactiveValues(doBox = FALSE)
-  observeEvent(input$goBofx, {
+  observeEvent(input$goBox, {
     # 0 will be coerced to FALSE
     # 1+ will be coerced to TRUE
     v10$doBox <- input$goBox
@@ -171,6 +187,68 @@ server <- function(input, output,session) {
     # 1+ will be coerced to TRUE
     v19$doWilcoxon <- input$goWilcoxon
   })
+  v20<-reactiveValues(gv1=0)
+  observeEvent(input$gv1,{
+    v20$gv1<-input$gv1
+  })
+  v21<-reactiveValues(gv2=0)
+  observeEvent(input$gv2,{
+    v21$gv2<-input$gv2
+  })
+  v22 <- reactiveValues(doChi = FALSE)
+  observeEvent(input$goChi, {
[…]

### 2020-07-27 shiny-app/webinar-aug20/ui.R

@@ -4,7 +4,7 @@ library(plotly)
 shinyUI(pageWithSidebar(
   
   headerPanel(
-    img(src = "logo.png", height = 75, width = 350),
+    img(src = "logo.png", height = 80, width = 350),
     # fix the adjust ratio 75/230
     # "Fast-Stats Web Tool",
     #           tags$head(tags$style(type="text/css", "label.radio { display: inline-block; }", ".radio input[type=\"radio\"] { float: none; }"),
@@ -54,22 +54,23 @@ shinyUI(pageWithSidebar(
     ),
     conditionalPanel(condition="input.tabs1=='Data visualization'",
                      # img(src = "logo.png", height = 75, width = 230),
-                     
-                     selectInput("plotType","Plot Type",choices =c("ViolinPlot"=0,"BoxPlot"=1)),
                      h4("Plot Option:"),
                      
+                     selectInput("plotType","Plot Type",choices =c("ViolinPlot"=0,"BoxPlot"=1)),
                      checkboxInput("addPoints", "Add data points", T),
                      
                      helpText("This option allows the user to add a scatterplot of the data where each dot corresponds to one observation (row) in the dataset."),
-                     
-                     h4("Data Visualization:"),
-                     
-                     fluidRow(
-                       column(6,radioButtons("xaxisGrp","Grouping Variables:", c("1"="1","2"="2"))),
-                       column(6,radioButtons("yaxisGrp","Quantity:", c("1"="1","2"="2")))
-                     ),
-                     
+                    
+                     # violinplot
                      conditionalPanel(condition="input.plotType==0",
+                                      
+                                      h4("Data Visualization:"),
+                                      
+                                      fluidRow(
+                                        column(6,radioButtons("xaxisGrp","Grouping Variables:", c("1"="1","2"="2"))),
+                                        column(6,radioButtons("yaxisGrp","Quantity:", c("1"="1","2"="2")))
+                                      ),
+                                      
                                       fluidRow(
                                         column(6, align="center", offset = 3,
                                                actionButton("goViolin", "Violin Plot"),
@@ -78,7 +79,15 @@ shinyUI(pageWithSidebar(
                                       ),
                                       helpText("Plot a numerical variable 'Quantity' by groups 'Group variable'. It is similar to the box plot but it also shows the distribution and spread of the data.")
                      ),
+                     # boxplot
                      conditionalPanel(condition="input.plotType==1",
+                                      
+                                      h4("Data Visualization:"),
+                                      
[…]

### 2020-07-28 shiny-app/webinar-aug20/server.R
fix a bunch of bugs; refine the UI; remove null output; fix the T-test bug

@@ -3,6 +3,7 @@ library(ggplot2)
 library(car)
 library(plotly)
 library(pracma)
+# library(shinyjs)
 # Define server logic to read selected file ----
 server <- function(input, output,session) {
   # #################################### 
@@ -25,7 +26,7 @@ server <- function(input, output,session) {
     cb_options <- list()
     cb_options[ dsnames] <- dsnames
     updateRadioButtons(session, "xaxisGrp",
-                       label = "Group Variable",
+                       label = "Grouping Variable",
                        choices = cb_options,
                        selected = "")
     updateRadioButtons(session, "yaxisGrp",
@@ -33,7 +34,7 @@ server <- function(input, output,session) {
                        choices = cb_options,
                        selected = "")
     updateRadioButtons(session, "groupVar",
-                       label = "Group Variable",
+                       label = "Grouping Variable",
                        choices = cb_options,
                        selected = "")
     updateRadioButtons(session, "quantity",
@@ -41,11 +42,11 @@ server <- function(input, output,session) {
                        choices = cb_options,
                        selected = "")
     updateRadioButtons(session, "gv1",
-                       label = "Group Variable 1",
+                       label = "Grouping Variable 1",
                        choices = cb_options,
                        selected = "")
     updateRadioButtons(session, "gv2",
-                       label = "Group Variable 2",
+                       label = "Grouping Variable 2",
                        choices = cb_options,
                        selected = "")
     updateRadioButtons(session, "gvBox",
@@ -66,6 +67,7 @@ server <- function(input, output,session) {
     # 0 will be coerced to FALSE
     # 1+ will be coerced to TRUE
     v$doViolinPlot <- input$goViolin
+    
   })
   
   v2 <- reactiveValues(doHist = FALSE)
@@ -89,6 +91,7 @@ server <- function(input, output,session) {
     # 0 will be coerced to FALSE
     # 1+ will be coerced to TRUE
     v4$doT <- input$goT
+    
   })
   
   v5<-reactiveValues(xv=0)
@@ -212,7 +215,20 @@ server <- function(input, output,session) {
   
   not_equalGV<-function(input1, input2){
     if(strcmp(input1, input2)){
-      showNotification("Please select different grouping variables!",duration=3,type = "error")
[…]

### 2020-07-28 shiny-app/webinar-aug20/ui.R

@@ -1,6 +1,6 @@
 library(shiny)
 library(plotly)
-
+  
 shinyUI(pageWithSidebar(
   
   headerPanel(
@@ -16,6 +16,7 @@ shinyUI(pageWithSidebar(
   ),
   
   sidebarPanel(
+    # useShinyjs(),
     conditionalPanel(condition="input.tabs1=='About'",
                      # img(src = "logo.png", height = 75, width = 230),
                      h4("Introduction")
@@ -59,15 +60,15 @@ shinyUI(pageWithSidebar(
                      selectInput("plotType","Plot Type",choices =c("ViolinPlot"=0,"BoxPlot"=1)),
                      checkboxInput("addPoints", "Add data points", T),
                      
-                     helpText("This option allows the user to add a scatterplot of the data where each dot corresponds to one observation (row) in the dataset."),
-                    
+                     HTML('<p style="color:#808080"> <b>Plot Option: </b> Add data points. This option allows the user to add a scatterplot of the data where each dot corresponds to one observation (row) in the dataset. </p>'),
+                     
                      # violinplot
                      conditionalPanel(condition="input.plotType==0",
                                       
                                       h4("Data Visualization:"),
                                       
                                       fluidRow(
-                                        column(6,radioButtons("xaxisGrp","Grouping Variables:", c("1"="1","2"="2"))),
+                                        column(6,radioButtons("xaxisGrp","Grouping Variable:", c("1"="1","2"="2"))),
                                         column(6,radioButtons("yaxisGrp","Quantity:", c("1"="1","2"="2")))
                                       ),
                                       
@@ -77,7 +78,7 @@ shinyUI(pageWithSidebar(
                                                tags$style(type='text/css', "#button { vertical-align: middle; height: 50px; width: 100%; font-size: 30px;}")
                                         ),
                                       ),
-                                      helpText("Plot a numerical variable 'Quantity' by groups 'Group variable'. It is similar to the box plot but it also shows the distribution and spread of the data.")
+                                      HTML('<p style="color:#808080"> <b>Violin Plot:</b> Plot a numerical variable ("Quantity") by groups ("Grouping variable"). It is similar to the box plot but it also shows the distribution and spread of the data. </p>'),
                      ),
                      # boxplot
                      conditionalPanel(condition="input.plotType==1",
@@ -85,7 +86,7 @@ shinyUI(pageWithSidebar(
                                       h4("Data Visualization:"),
                                       
[…]

### 2020-07-29 shiny-app/webinar-aug20/server.R
figure out that input.actionbutton val won't work for removing null output;change all "grouping" into "group"

@@ -26,7 +26,7 @@ server <- function(input, output,session) {
     cb_options <- list()
     cb_options[ dsnames] <- dsnames
     updateRadioButtons(session, "xaxisGrp",
-                       label = "Grouping Variable",
+                       label = "Group Variable",
                        choices = cb_options,
                        selected = "")
     updateRadioButtons(session, "yaxisGrp",
@@ -34,7 +34,7 @@ server <- function(input, output,session) {
                        choices = cb_options,
                        selected = "")
     updateRadioButtons(session, "groupVar",
-                       label = "Grouping Variable",
+                       label = "Group Variable",
                        choices = cb_options,
                        selected = "")
     updateRadioButtons(session, "quantity",
@@ -42,15 +42,15 @@ server <- function(input, output,session) {
                        choices = cb_options,
                        selected = "")
     updateRadioButtons(session, "gv1",
-                       label = "Grouping Variable 1",
+                       label = "Group Variable 1",
                        choices = cb_options,
                        selected = "")
     updateRadioButtons(session, "gv2",
-                       label = "Grouping Variable 2",
+                       label = "Group Variable 2",
                        choices = cb_options,
                        selected = "")
     updateRadioButtons(session, "gvBox",
-                       label = "Grouping Variable",
+                       label = "Group Variable",
                        choices = cb_options,
                        selected = "")
     updateRadioButtons(session, "qBox",
@@ -150,17 +150,17 @@ server <- function(input, output,session) {
   })
   
   v14 <- reactiveValues(sel1 =0)
-  observeEvent(input$group1, {
+  observeEvent(input$Group1, {
     # 0 will be coerced to FALSE
     # 1+ will be coerced to TRUE
-    v14$sel1<- input$group1
+    v14$sel1<- input$Group1
   })
   
   v15 <- reactiveValues(sel2 =0)
-  observeEvent(input$group2, {
+  observeEvent(input$Group2, {
     # 0 will be coerced to FALSE
     # 1+ will be coerced to TRUE
-    v15$sel2<- input$group2
+    v15$sel2<- input$Group2
   })
   
   v16 <- reactiveValues(doEqualVar=F)
[…]

### 2020-07-29 shiny-app/webinar-aug20/ui.R

@@ -68,7 +68,7 @@ shinyUI(pageWithSidebar(
                                       h4("Data Visualization:"),
                                       
                                       fluidRow(
-                                        column(6,radioButtons("xaxisGrp","Grouping Variable:", c("1"="1","2"="2"))),
+                                        column(6,radioButtons("xaxisGrp","Group Variable:", c("1"="1","2"="2"))),
                                         column(6,radioButtons("yaxisGrp","Quantity:", c("1"="1","2"="2")))
                                       ),
                                       
@@ -78,7 +78,7 @@ shinyUI(pageWithSidebar(
                                                tags$style(type='text/css', "#button { vertical-align: middle; height: 50px; width: 100%; font-size: 30px;}")
                                         ),
                                       ),
-                                      HTML('<p style="color:#808080"> <b>Violin Plot:</b> Plot a numerical variable ("Quantity") by groups ("Grouping variable"). It is similar to the box plot but it also shows the distribution and spread of the data. </p>'),
+                                      HTML('<p style="color:#808080"> <b>Violin Plot:</b> Plot a numerical variable ("Quantity") by groups ("Group variable"). It is similar to the box plot but it also shows the distribution and spread of the data. </p>'),
                      ),
                      # boxplot
                      conditionalPanel(condition="input.plotType==1",
@@ -86,7 +86,7 @@ shinyUI(pageWithSidebar(
                                       h4("Data Visualization:"),
                                       
                                       fluidRow(
-                                        column(6,radioButtons("gvBox","Grouping Variable:", c("1"="1","2"="2"))),
+                                        column(6,radioButtons("gvBox","Group Variable:", c("1"="1","2"="2"))),
                                         column(6,radioButtons("qBox","Quantity:", c("1"="1","2"="2")))
                                       ),
                                       fluidRow(
@@ -95,7 +95,7 @@ shinyUI(pageWithSidebar(
                                                tags$style(type='text/css', "#button { vertical-align: middle; height: 50px; width: 100%; font-size: 30px;}")
                                         )
                                       ),
-                                      HTML('<p style="color:#808080"> <b>Box Plot:</b> Plot a numerical variable ("Quantity") by groups ("Grouping variable"). Solid black line in the box represents the median, and the upper and lower edges of the box represent the 3rd and 1st quartiles respectively. </p>'),
[…]

### 2020-07-29 shiny-app/webinar-aug20/server.R
removing null output successfully

@@ -150,17 +150,17 @@ server <- function(input, output,session) {
   })
   
   v14 <- reactiveValues(sel1 =0)
-  observeEvent(input$Group1, {
+  observeEvent(input$group1, {
     # 0 will be coerced to FALSE
     # 1+ will be coerced to TRUE
-    v14$sel1<- input$Group1
+    v14$sel1<- input$group1
   })
   
   v15 <- reactiveValues(sel2 =0)
-  observeEvent(input$Group2, {
+  observeEvent(input$group2, {
     # 0 will be coerced to FALSE
     # 1+ will be coerced to TRUE
-    v15$sel2<- input$Group2
+    v15$sel2<- input$group2
   })
   
   v16 <- reactiveValues(doEqualVar=F)
@@ -216,7 +216,7 @@ server <- function(input, output,session) {
   not_equalGV<-function(input1, input2){
     if(strcmp(input1, input2)){
       showNotification("Please select different 'Group Variables'!",duration=3,type = "error")
-      # "Please select different Group variables!"
+      # "Please select different grouping variables!"
     }else if(input1=="" || input2=="")
     {
       F
@@ -228,8 +228,8 @@ server <- function(input, output,session) {
   
   not_equalGV2<-function(input1, input2){
     if(strcmp(input1, input2)){
-      showNotification("Please select different Group variables for 'Group one' and 'Group two'!",duration=3,type = "error")
-      # "Please select different Group variables!"
+      showNotification("Please select different group variables for 'Group one' and 'Group two'!",duration=3,type = "error")
+      # "Please select different grouping variables!"
     }else if(input1=="" || input2=="")
     {
       F
@@ -243,7 +243,7 @@ server <- function(input, output,session) {
     # check to see the data type of a specific col in data frame
     if(class(df[[input]])=="numeric"||class(df[[input]])=="integer"||class(df[[input]])=="complex"){
       showNotification( "Please select a categorical variable to be the 'Group Variable'!",duration=3,type = "error")
-      # "Please select a categorical variable to be the Group variable!"
+      # "Please select a categorical variable to be the grouping variable!"
     }else if(df==""||input==""){
       F
     }
@@ -256,7 +256,7 @@ server <- function(input, output,session) {
[…]

### 2020-07-29 shiny-app/webinar-aug20/ui.R

@@ -1,6 +1,6 @@
 library(shiny)
 library(plotly)
-  
+
 shinyUI(pageWithSidebar(
   
   headerPanel(
@@ -99,137 +99,137 @@ shinyUI(pageWithSidebar(
                                       
                      ),
     ),
-                     conditionalPanel(condition="input.tabs1=='Data analysis'",
-                                      # img(src = "logo.png", height = 75, width = 230),
+    conditionalPanel(condition="input.tabs1=='Data analysis'",
+                     # img(src = "logo.png", height = 75, width = 230),
+                     
+                     # h4("Data Analysis:"),
+                     
+                     selectInput("testType","Test Type",choices =c("Chi-Square Test"=0,"T-Test"=1)),
+                     # t-test
+                     conditionalPanel(condition="input.testType==1",
+                                      h4("Data Analysis Option:"),
                                       
-                                      # h4("Data Analysis:"),
+                                      checkboxInput("isEqualVar", "Equal Variance", F),
                                       
-                                      selectInput("testType","Test Type",choices =c("Chi-Square Test"=0,"T-Test"=1)),
-                                      # t-test
-                                      conditionalPanel(condition="input.testType==1",
-                                                       h4("Data Analysis Option:"),
-                                                       
-                                                       checkboxInput("isEqualVar", "Equal Variance", F),
-                                                       
-                                                       HTML('<p style="color:#808080"> <b> Equal variance:</b> The standard t test assumes equal variances on the two groups. If the user checks this option, the standard t test is run, but if the user unchecks this option, then the Welch t test is run instead (that does not assume equal variances). </p>'),
-                                                       fluidRow(
-                                                         column(6,radioButtons("groupVar","Group Variable:", c("1"="UnSpecified_Value","2"="UnSpecified_Value"))),
-                                                         column(6,radioButtons("quantity","Quantity:", c("1"="UnSpecified_Value","2"="UnSpecified_Value")))
-                                                       ),
-                                                       uiOutput("sel1"),
-                                                       uiOutput("sel2"),
-                                                       HTML('<p style="color:#808080"> <b>T test:</b> Statistical test of the null hypothesis of equality of means of a numerical variable ("Quantity") on two groups ("Group variable"). If the selected group variable has more than two categories, the user will select the two groups to compare. </p>'),
-                                                       fluidRow(
-                                                         column(6, align="center", offset = 3,
[…]

### 2020-07-29 cotyledon-data-analysis.Rmd
fixed typo

@@
 dat2 = data.frame(generation=dat$Generation)
-dat2$cots = "dycot"
+dat2$cots = "dicot"
 dat2$cots[dat$Cotyledons > 2] = "polycot"

### 2020-07-31 shiny-app/webinar-aug20/ui.R
Add mosaic plot (not in plotly form, need to fix it later); add chi-squ feature; fix minor bugs

@@ -57,13 +57,33 @@ shinyUI(pageWithSidebar(
                      # img(src = "logo.png", height = 75, width = 230),
                      h4("Plot Option:"),
                      
-                     selectInput("plotType","Plot Type",choices =c("ViolinPlot"=0,"BoxPlot"=1)),
-                     checkboxInput("addPoints", "Add data points", T),
+                     selectInput("plotType","Plot Type",choices =c("MosaicPlot"=0,"ViolinPlot"=1,"BoxPlot"=2)),
                      
-                     HTML('<p style="color:#808080"> <b>Plot Option: </b> Add data points. This option allows the user to add a scatterplot of the data where each dot corresponds to one observation (row) in the dataset. </p>'),
+                     # mosaic plot
+                     conditionalPanel(condition="input.plotType==0",
+                                      
+                                      h4("Data Visualization:"),
+                                      
+                                      fluidRow(
+                                        column(6,radioButtons("gvMosaic1","Group Variable 1:", c("1"="1","2"="2"))),
+                                        column(6,radioButtons("gvMosaic2","Group Variable 2:", c("1"="1","2"="2")))
+                                      ),
+                                      
+                                      fluidRow(
+                                        column(6, align="center", offset = 3,
+                                               actionButton("goMosaic", "Mosaic Plot"),
+                                               tags$style(type='text/css', "#button { vertical-align: middle; height: 50px; width: 100%; font-size: 30px;}")
+                                        ),
+                                      ),
+                                      HTML('<p style="color:#808080"> <b>Mosaic plot:</b> Plot to visualize contigency tables of frequencies among categorical variables </p>'),
+                     ),
                      
                      # violinplot
-                     conditionalPanel(condition="input.plotType==0",
+                     conditionalPanel(condition="input.plotType==1",
+                                      
+                                      checkboxInput("addPoints", "Add data points", T),
+                                      
+                                      HTML('<p style="color:#808080"> <b>Add data points: </b> This option allows the user to add a scatterplot of the data where each dot corresponds to one observation (row) in the dataset. </p>'),
                                       
                                       h4("Data Visualization:"),
                                       
@@ -81,7 +101,11 @@ shinyUI(pageWithSidebar(
[…]

### 2020-07-31 cotyledon-data-analysis.Rmd
added mosaic

@@
 mosaicplot(dt, shade = TRUE, las=2, main="Cotelydons")
+
+library(ggmosaic)
+ggplot(data = dat2) +
+  geom_mosaic(aes(x = product(cots), fill=generation))
 ```

### 2020-07-31 shiny-app/webinar-aug20/server.R
add titles to three plots; add new lines to chi-sq test

@@ -395,7 +395,7 @@ server <- function(input, output,session) {
     y_val<-unlist(subset(df, select=c(v6$yv)))
     
     # options(repr.plot.width=v14$doWidthVal, repr.plot.height=v15$doHeightVal)
-    plot<-ggplot(df, aes(x=x_val, y=y_val, fill=x_val))+geom_violin(alpha=0.5)+xlab(v5$xv)+ylab(v6$yv)+labs(fill=v5$xv)+
+    plot<-ggplot(df, aes(x=x_val, y=y_val, fill=x_val))+geom_violin(alpha=0.5)+xlab(v5$xv)+ylab(v6$yv)+labs(fill=v5$xv)+ggtitle("Violin Plot")+
       ylim(c(1,6))+
       theme(
         plot.title = element_text(hjust=0.5, size=rel(1.8)),
@@ -451,16 +451,14 @@ server <- function(input, output,session) {
     )
     g_val1<-unlist(subset(df, select=c(v25$gvMosaic1)))
     g_val2<-unlist(subset(df, select=c(v26$gvMosaic2)))
-    dt = table(g_val1,  g_val2)   
-    mosaicplot(dt, shade = TRUE, las=2,xlab=v25$gvMosaic1,ylab=v26$gvMosaic2)
-    # plot<-ggplot(dt)
-    # ggplotly(plot)
-    # print(g_val2)
-    # # plot<-ggplot(df) +
-    # #   geom_mosaic(aes(weight=g_val2,x = product(g_val1), 
-    # #                   fill = factor(g_val1),na.rm=T)) 
-    # df<-data(df)
-    # plot<-as.ggplot(~mosaic(df))
+    dt = table(g_val1,  g_val2)
+    mosaicplot(dt, shade = TRUE, las=2,xlab=v25$gvMosaic1,ylab=v26$gvMosaic2,main="Mosaic Plot")
+    
+# ################################################
+     # ggplot(df) +
+     #  geom_mosaic(aes(x=product(gval1), fill=gval2),data=df)+xlab(v25$gvMosaic1)+ylab(v26$gvMosaic2)+ggtitle("Mosaic Plot")
+    # plot<-ggplot(df) +
+    #   geom_mosaic(aes(x=product(cots), fill=generation))+xlab(v25$gvMosaic1)+ylab(v26$gvMosaic2)
     # ggplotly(plot)
   })
   
@@ -614,9 +612,9 @@ server <- function(input, output,session) {
[…]

### 2020-08-01 shiny-app/webinar-aug20/ui.R
finally enables mosaic work together with plotly

@@ -216,7 +216,7 @@ shinyUI(pageWithSidebar(
       # plot tab
       tabPanel("Data visualization", 
                conditionalPanel(condition = "input.plotType==0",
-                                plotOutput("mosaicPlot") ),
+                                plotlyOutput("mosaicPlot") ),
                conditionalPanel(condition = "input.plotType==1",
                                 plotlyOutput("violinPlot") ),
                conditionalPanel(condition = "input.plotType==2",

### 2020-09-17 shiny-app/webinar-aug20/server.R
color palettes added

@@ -7,6 +7,8 @@ library(graphics)
 library(ggplotify)
 library(thatssorandom)
 library(rsconnect)
+# library(viridis)
+library(RColorBrewer)
 # library(shinyjs)
 # Define server logic to read selected file ----
 server <- function(input, output,session) {
@@ -287,6 +289,10 @@ server <- function(input, output,session) {
   v35<-reactiveValues(pointSizeScatter=0)
   observeEvent(input$pointSizeScatter,{
     v35$pointSizeScatter<-input$pointSizeScatter
+  })
+  v36<-reactiveValues(colorScatter=0)
+  observeEvent(input$colorScatter,{
+    v36$colorScatter<-input$colorScatter
   })
   
   not_equalGV<-function(input1, input2){
@@ -613,7 +619,7 @@ server <- function(input, output,session) {
     
     p<-ggplot(df, aes(y = y_val, x = x_val, fill = x_val)) +
       xlab(v29$gvScatter)+labs(fill=v29$gvScatter)+ylab(v30$qScatter)+
-      geom_jitter(pch = v34$pointShapeScatter, alpha=v33$transScatter/100, width=0.2,size=v35$pointSizeScatter/100)+
+      geom_jitter(pch = v34$pointShapeScatter, alpha=v33$transScatter/100, width=0.2,size=v35$pointSizeScatter)+
       theme(
         plot.title = element_text(hjust=0.5, size=rel(1.8)),
         axis.title.x = element_text(size=rel(1.8)),
@@ -623,7 +629,10 @@ server <- function(input, output,session) {
         panel.background = element_blank(),
         text=element_text(size=8),
         axis.line = element_line(colour = "grey")##,
-      ) 
+      )
+    # brewer.pal(n = 8, name = "Spectral")
+    p<-p+ scale_fill_brewer(palette = v36$colorScatter)
+    # A character string indicating the colormap option to use. Four options are available: "magma" (or "A"), "inferno" (or "B"), "plasma" (or "C"), "viridis" (or "D", the default option) and "cividis" (or "E").
     ggplotly(p)
     
     

### 2020-09-17 shiny-app/webinar-aug20/ui.R

@@ -197,13 +197,15 @@ shinyUI(dashboardPage(
                              
                              h4("Data Visualization:"),
                              
+                             selectInput("colorScatter","Color",choices =c("Blue+Purple"="BuPu","Dark Color"="Dark2","Orange+Red"="OrRd","Yellow+Green+Blue"="YlGnbu","Grey"="Greys","Paired"="Paired","Red+Blue"="RdBu","Purple+Red"="PuRd","Blue"="Blues","Red"="Reds")),
+                             
                              sliderInput("transScatter", "Transparency:",
                                          min = 0, max = 100,
-                                         value = 30),
+                                         value = 50),
                              
                              sliderInput("pointSizeScatter", "Point Size:",
-                                         min = 0, max = 100,
-                                         value = 50),
+                                         min = 0, max = 10,
+                                         value = 2),
                              
                              selectInput("pointShapeScatter","Point Shape",choices =c("Solid Circle"=19,"Bullet"=20,"Filled Circle"=21,"Filled Square"=22,"Filled Diamond"=23,"Filled Triangle Point-Up"=24, "Filled Triangle Point-down"=25)),
                              
@@ -280,6 +282,7 @@ shinyUI(dashboardPage(
                                )
                              ),
                              HTML('<p style="color:#808080"> <b>How to interpret the result?</b> If the p-value is less than 0.05, we reject the null hypothesis of equality of means. The confidence interval represents the interval for the difference of means. </p>'),
+                             HTML('<p style="color:#808080"> <b>Assumptions of t test:</b> The t test assumes normality, equal variance and independence. Take a look at <a href="https://wolfganghuber.shinyapps.io/t-test-normality-and-independence/" target="_blank">this link</a> that illustrate how important normality and independence are in the t test results. </p>'),
             ),
             # chi-square
             conditionalPanel(condition="input.testType==0",

### 2020-09-17 shiny-app/webinar-aug20/server.R
increase contrasting color effect

@@ -7,7 +7,7 @@ library(graphics)
 library(ggplotify)
 library(thatssorandom)
 library(rsconnect)
-# library(viridis)
+library(viridis)
 library(RColorBrewer)
 # library(shinyjs)
 # Define server logic to read selected file ----
@@ -630,9 +630,17 @@ server <- function(input, output,session) {
         text=element_text(size=8),
         axis.line = element_line(colour = "grey")##,
       )
-    # brewer.pal(n = 8, name = "Spectral")
-    p<-p+ scale_fill_brewer(palette = v36$colorScatter)
-    # A character string indicating the colormap option to use. Four options are available: "magma" (or "A"), "inferno" (or "B"), "plasma" (or "C"), "viridis" (or "D", the default option) and "cividis" (or "E").
+    group_list=unlist(subset(df,select=c(v29$gvScatter)))
+    colorCount = length(unique(unlist(group_list,use.names=F)))   #8, an arbitrary number
+      # length(unique(subset(df, select=c(v29$gvScatter))))
+    getPalette <- colorRampPalette(brewer.pal(8, v36$colorScatter),bias=5)(colorCount)
+    p<-p+scale_fill_manual(values= getPalette)
+    # p<-p+ scale_fill_brewer(palette = v36$colorScatter)+scale_color_brewer(palette = v36$colorScatter)
+    # p<-p+ scale_fill_brewer(palette = v36$colorScatter,direction=-1)+scale_color_brewer(palette = v36$colorScatter,direction=-1)
+   
+    # p<-p+scale_color_viridis(discrete = T, option = v36$colorScatter)+
+    #   scale_fill_viridis(discrete = T) 
+   
     ggplotly(p)
     
     

### 2020-09-17 shiny-app/webinar-aug20/ui.R

@@ -197,11 +197,14 @@ shinyUI(dashboardPage(
                              
                              h4("Data Visualization:"),
                              
-                             selectInput("colorScatter","Color",choices =c("Blue+Purple"="BuPu","Dark Color"="Dark2","Orange+Red"="OrRd","Yellow+Green+Blue"="YlGnbu","Grey"="Greys","Paired"="Paired","Red+Blue"="RdBu","Purple+Red"="PuRd","Blue"="Blues","Red"="Reds")),
+                             selectInput("colorScatter","Color",choices = c("Blue+Purple"="BuPu","Dark Color"="Dark2","Orange+Red"="OrRd","Yellow+Green+Blue"="YlGnBu","Accent"="Accent","Paired"="Paired","Red+Blue"="RdYlBu","Purple+Red"="PuRd","Set2"="Set2","Purple+Green"="PRGn")
+                             ),
                              
+                             # c("viridis"="viridis","magma"="magma","inferno"="inferno","plasma"="plasma","cividis"="cividis")
+                            
                              sliderInput("transScatter", "Transparency:",
-                                         min = 0, max = 100,
-                                         value = 50),
+                                         min = 30, max = 100,
+                                         value = 65),
                              
                              sliderInput("pointSizeScatter", "Point Size:",
                                          min = 0, max = 10,

### 2020-09-20 shiny-app/webinar-aug20/server.R
add some comments and the color palette img

@@ -633,7 +633,9 @@ server <- function(input, output,session) {
     group_list=unlist(subset(df,select=c(v29$gvScatter)))
     colorCount = length(unique(unlist(group_list,use.names=F)))   #8, an arbitrary number
       # length(unique(subset(df, select=c(v29$gvScatter))))
-    getPalette <- colorRampPalette(brewer.pal(8, v36$colorScatter),bias=5)(colorCount)
+    getPalette <- colorRampPalette(brewer.pal(8, v36$colorScatter),bias=2.5)(colorCount)
+    # FIXME LATER:
+    # bias value needs to be tested to get the best level change within color palette
     p<-p+scale_fill_manual(values= getPalette)
     # p<-p+ scale_fill_brewer(palette = v36$colorScatter)+scale_color_brewer(palette = v36$colorScatter)
     # p<-p+ scale_fill_brewer(palette = v36$colorScatter,direction=-1)+scale_color_brewer(palette = v36$colorScatter,direction=-1)

### 2020-09-20 ecosystem-data-analysis.Rmd
ecosystem data

---
title: "WI FastPlants: Ecosystem in plant indicators"
author: "Claudia Solis-Lemus"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(car)
```

# Initial data: `ecosystem-data.csv`

### Reading the data

Reading in the data (had to remove manually the '#' sign from the column and save as csv file; also I removed the rows of "dead plant" and the text with the description of variables. All in the original file: `102314-Phenotypic-Plasticity-Data.xlsx`):
```{r}
dat = read.table("data/ecosystem-data.csv", sep=",",header=TRUE)
str(dat)
```

The data has 9 columns:
```{r}
head(dat)
```

Some of the variables that need to be numbers, as actually read as character due to some cell values being words.
```{r}
dat[dat$Flowers == "dropsy",]
dat2 = dat[dat$Flowers != "dropsy",]
dat2$Flowers = as.numeric(dat2$Flowers)
str(dat2)
```

## Data visualization

We will focus on visualizations for a dependent variable that is numeric (e.g. height, cot width, flowers, bracts) with respect to an independent variable that is also numeric (osmocote).

We will do the plots only for X=osmocote, Y=height, but these same plots can be done for Y=CotWidth or Y=Flowers or Y=Bracts, or any other numerical variable in future datasets.

```{r, echo=FALSE}
p2 = ggplot(dat2, aes(x = Osmocote, y = Height, fill = Osmocote)) +
  geom_point(pch = 21, alpha=0.3, position = position_jitterdodge(jitter.height=0.075, jitter.width=0.1))+
  theme(
        plot.title = element_text(hjust=0.5, size=rel(1.8)),
        axis.title.x = element_text(size=rel(1.8)),
        axis.title.y = element_text(size=rel(1.8), angle=90, vjust=0.5, hjust=0.5),
        axis.text.x = element_text(colour="grey", size=rel(1.5), angle=0, hjust=.5, vjust=.5, face="plain"),
        axis.text.y = element_text(colour="grey", size=rel(1.5), angle=0, hjust=.5, vjust=.5, face="plain"),
        panel.background = element_blank(),
        axis.line = element_line(colour = "grey")##,
        )
p2
```

We want to give people the option to add a linear regression line:
```{r, echo=FALSE}
p2 = ggplot(dat2, aes(x = Osmocote, y = Height, fill = Osmocote)) +
  geom_point(pch = 21, alpha=0.3, position = position_jitterdodge(jitter.height=0.075, jitter.width=0.1))+
  #ylim(c(1,6))+
  theme(
        plot.title = element_text(hjust=0.5, size=rel(1.8)),
        axis.title.x = element_text(size=rel(1.8)),
        axis.title.y = element_text(size=rel(1.8), angle=90, vjust=0.5, hjust=0.5),
        axis.text.x = element_text(colour="grey", size=rel(1.5), angle=0, hjust=.5, vjust=.5, face="plain"),
        axis.text.y = element_text(colour="grey", size=rel(1.5), angle=0, hjust=.5, vjust=.5, face="plain"),
        panel.background = element_blank(),
        axis.line = element_line(colour = "grey")##,
        )
p2
```

### 2020-09-21 shiny-app/webinar-aug20/server.R
box and densities get fixed

@@ -294,6 +294,75 @@ server <- function(input, output,session) {
   observeEvent(input$colorScatter,{
     v36$colorScatter<-input$colorScatter
   })
+  ##
+  v37<-reactiveValues(transMosaic=0)
+  observeEvent(input$transMosaic,{
+    v37$transMosaic<-input$transMosaic
+  })
+  v38<-reactiveValues(pointShapeMosaic=0)
+  observeEvent(input$pointShapeMosaic,{
+    v38$pointShapeMosaic<-input$pointShapeMosaic
+  })
+  v39<-reactiveValues(pointSizeMosaic=0)
+  observeEvent(input$pointSizeScatter,{
+    v39$pointSizeMosaic<-input$pointSizeMosaic
+  })
+  v40<-reactiveValues(colorMosaic=0)
+  observeEvent(input$colorMosaic,{
+    v40$colorMosaic<-input$colorMosaic
+  })
+  ##
+  v41<-reactiveValues(transViolin=0)
+  observeEvent(input$transViolin,{
+    v41$transViolin<-input$transViolin
+  })
+  v42<-reactiveValues(pointShapeViolin=0)
+  observeEvent(input$pointShapeViolin,{
+    v42$pointShapeViolin<-input$pointShapeViolin
+  })
+  v43<-reactiveValues(pointSizeViolin=0)
+  observeEvent(input$pointSizeViolin,{
+    v43$pointSizeViolin<-input$pointSizeViolin
+  })
+  v44<-reactiveValues(colorViolin=0)
+  observeEvent(input$colorViolin,{
+    v44$colorViolin<-input$colorViolin
+  })
+  ##box
+  v45<-reactiveValues(transBox=0)
+  observeEvent(input$transBox,{
+    v45$transBox<-input$transBox
+  })
+  v46<-reactiveValues(pointShapeBox=0)
+  observeEvent(input$pointShapeBox,{
+    v46$pointShapeBox<-input$pointShapeBox
+  })
+  v47<-reactiveValues(pointSizeBox=0)
+  observeEvent(input$pointSizeBox,{
+    v47$pointSizeBox<-input$pointSizeBox
+  })
+  v48<-reactiveValues(colorBox=0)
+  observeEvent(input$colorBox,{
+    v48$colorBox<-input$colorBox
+  })
+  ##densities
+  v49<-reactiveValues(transDensities=0)
+  observeEvent(input$transDensities,{
+    v49$transDensities<-input$transDensities
[…]

### 2020-09-21 shiny-app/webinar-aug20/ui.R

@@ -31,7 +31,7 @@ shinyUI(dashboardPage(
       menuItem("Data Upload", tabName = "dataUpload"),
       menuItem("Data Visualization", tabName = "dataVisualization"),
       menuItem("Data Analysis", tabName = "dataAnalysis"),
-      menuItem("FAQ", tabName = "faq")
+      menuItem("FAQ", tabName = "FAQ")
     )
   ),
   
@@ -71,7 +71,7 @@ shinyUI(dashboardPage(
               box( width=12, status="primary",solidHeader = T,
                    title = "Welcome to the WI Fast Stats app!",
                    
-                   HTML( 'This is the <a href="https://github.com/crsl4/fast-stats" target="_blank">open-source</a> publicly available web app to analyze data from <a href="https://fastplants.org/" target="_blank">WI Fast Plants</a>.'))
+                   h4(HTML( 'This is the <a href="https://github.com/crsl4/fast-stats" target="_blank">open-source</a> publicly available web app to analyze data from <a href="https://fastplants.org/" target="_blank">WI Fast Plants</a>.')))
               
       ),
       
@@ -149,6 +149,13 @@ shinyUI(dashboardPage(
                                column(6,radioButtons("gvMosaic1","Group Variable 1:", c("1"="1","2"="2"))),
                                column(6,radioButtons("gvMosaic2","Group Variable 2:", c("1"="1","2"="2")))
                              ),
+                             selectInput("colorMosaic","Color",choices = c("Blue+Purple"="BuPu","Dark Color"="Dark2","Orange+Red"="OrRd","Yellow+Green+Blue"="YlGnBu","Accent"="Accent","Paired"="Paired","Red+Blue"="RdYlBu","Purple+Red"="PuRd","Set2"="Set2","Purple+Green"="PRGn")
+                             ),
+                             HTML('<p style="color:#808080"> <b>Notice:</b> Click the <b>FAQ</b> tab in the left sidebar to see the detailed color palettes</p>'),
+                             
+                             sliderInput("transMosaic", "Transparency:",
+                                         min = 30, max = 100,
+                                         value = 65),
                              
                              fluidRow(
                                column(6, align="center", offset = 3,
@@ -173,6 +180,20 @@ shinyUI(dashboardPage(
                                column(6,radioButtons("yaxisGrp","Quantity:", c("1"="1","2"="2")))
                              ),
                              
[…]

### 2020-09-21 shiny-app/webinar-aug20/server.R
violin fixed; only left mosaic transparency

@@ -515,7 +515,7 @@ server <- function(input, output,session) {
     y_val<-unlist(subset(df, select=c(v6$yv)))
     
     # options(repr.plot.width=v14$doWidthVal, repr.plot.height=v15$doHeightVal)
-    plot<-ggplot(df, aes(x=x_val, y=y_val, fill=x_val))+geom_violin(alpha=0.5)+xlab(v5$xv)+ylab(v6$yv)+labs(fill=v5$xv)+ggtitle("Violin Plot")+
+    plot<-ggplot(df, aes(x=x_val, y=y_val, fill=x_val))+geom_violin(alpha=v41$transViolin/100)+xlab(v5$xv)+ylab(v6$yv)+labs(fill=v5$xv)+ggtitle("Violin Plot")+
       ylim(c(1,6))+
       theme(
         plot.title = element_text(hjust=0.5, size=rel(1.8)),
@@ -528,8 +528,13 @@ server <- function(input, output,session) {
         axis.line = element_line(colour = "grey")##,
       )  
     if(v12$doAddPoints==T){
-      plot<-plot+geom_point(pch = 21, alpha=0.3, position = position_jitterdodge(jitter.height=0.05, jitter.width=2.5),size=0.5)
+      plot<-plot+geom_point(pch = v42$pointShapeViolin, alpha=v41$transViolin/100, position = position_jitterdodge(jitter.height=0.05, jitter.width=2.5),size=v43$pointSizeViolin)
+      
     }
+    group_list<-unlist(subset(df,select=c(v5$xv)))
+    colorCount=length(unique(unlist(group_list,use.names = F)))
+    getPalette<-colorRampPalette(brewer.pal(8,v44$colorViolin),bias=2.5)(colorCount)
+    plot<-plot+scale_fill_manual(values=getPalette)
     ggplotly(plot)
   })
   
@@ -579,6 +584,15 @@ server <- function(input, output,session) {
         plot.title = element_text(hjust = 0.5),
         text=element_text(size=9),
       )
+    group_list=unlist(subset(df,select=c(v25$gvMosaic1)))
+    colorCount = length(unique(unlist(group_list,use.names=F)))
+    group_list=unlist(subset(df,select=c(v26$gvMosaic2)))
+    colorCount = max(colorCount,length(unique(unlist(group_list,use.names=F))) )  
+    getPalette <- colorRampPalette(brewer.pal(8, v40$colorMosaic),bias=2.5)(colorCount)
+    # FIXME LATER:
[…]

### 2020-10-07 shiny-app/webinar-dec20/server.R
eco-system webapp initial commit

@@ -1,0 +1,1011 @@
+library(data.table)
+library(ggplot2)
+library(car)
+library(plotly)
+library(pracma)
+library(graphics)
+library(ggplotify)
+library(thatssorandom)
+library(rsconnect)
+library(viridis)
+library(RColorBrewer)
+# library(shinyjs)
+# Define server logic to read selected file ----
+server <- function(input, output,session) {
+  # #################################### 
+  dsnames<-c() #a vector to store col names
+  
+  data_set <- reactive({
+    req(input$file1)
+    inFile <- input$file1
+    read.csv(inFile$datapath, header=input$header, 
+             sep=input$sep)
+    
+  })
+  # dsnames0<-c() #a vector to store col names
+  # 
+  # data_set0 <- reactive({
+  #   read.csv("ecosystem-data.csv", header=T, 
+  #            sep=",")
+  # })
+  observe({
+    if(input$fileType=="uploadFile"){
+      dsnames <- names(data_set())
+      cb_options <- list()
+      cb_options[ dsnames] <- dsnames
+      updateRadioButtons(session, "xaxisGrp",
+                         label = "Group Variable",
+                         choices = cb_options,
+                         selected = "")
+      updateRadioButtons(session, "yaxisGrp",
+                         label = "Quantity",
+                         choices = cb_options,
+                         selected = "")
+      updateRadioButtons(session, "groupVar",
+                         label = "Group Variable",
+                         choices = cb_options,
+                         selected = "")
+      updateRadioButtons(session, "quantity",
+                         label = "Quantity",
+                         choices = cb_options,
+                         selected = "")
+      updateRadioButtons(session, "gv1",
+                         label = "Group Variable 1",
+                         choices = cb_options,
+                         selected = "")
+      updateRadioButtons(session, "gv2",
+                         label = "Group Variable 2",
+                         choices = cb_options,
+                         selected = "")
+      updateRadioButtons(session, "gvBox",
+                         label = "Group Variable",
+                         choices = cb_options,
+                         selected = "")
+      updateRadioButtons(session, "qBox",
[…]

### 2020-10-07 shiny-app/webinar-dec20/ui.R

@@ -1,0 +1,339 @@
+library(shiny)
+library(plotly)
+library(shinydashboard)
+# library(dashboardthemes)
+
+shinyUI(dashboardPage(
+  skin="black",
+ 
+  dashboardHeader(
+    # theme="blue_gradient",
+    
+    title="Fast-Stats Web Tool",
+    tags$li(img(src='v_line.png',width=20,height =50),class="dropdown"),
+    tags$li(img(src='logo.png',width=175,height =50),class="dropdown"),
+    tags$li(img(src='v_line.png',width=20,height =50),class="dropdown"),
+    tags$li(img(src='WIDSimpleAcronym.png',width=80,height =50),class="dropdown"),
+    tags$li(img(src='v_line.png',width=20,height =50),class="dropdown"),
+    tags$li(img(src='uw-logo-flush-web.png',width=150,height =50),class="dropdown"),
+    tags$li(img(src='v_line.png',width=20,height =50),class="dropdown")
+    
+    # fix the adjust ratio 75/230
+    # "Fast-Stats Web Tool",
+    #           tags$head(tags$style(type="text/css", "label.radio { display: inline-block; }", ".radio input[type=\"radio\"] { float: none; }"),
+    #                     tags$style(type="text/css", "select { max-width: 200px; }"),
+    #                     tags$style(type="text/css", "textarea { max-width: 185px; }"),
+    #                     tags$style(type="text/css", ".jslider { max-width: 200px; }"),
+    #                     tags$style(type='text/css', ".well { max-width: 330px; }"),
+    #                     tags$style(type='text/css', ".span4 { max-width: 330px; }")) 
+  ),
+  
+  dashboardSidebar(
+    sidebarMenu(
+      menuItem("About", tabName = "about"),
+      menuItem("Data Upload", tabName = "dataUpload"),
+      menuItem("Data Visualization", tabName = "dataVisualization"),
+      menuItem("FAQ", tabName = "FAQ")
+    )
+  ),
+  
+  dashboardBody(
+    # for internal links
+    tags$script(HTML("
+        var openTab = function(tabName){
+          $('a', $('.sidebar')).each(function() {
+            if(this.getAttribute('data-value') == tabName) {
+              this.click()
+            };
+          });
+        }
+      ")),
+    tags$head(
+      tags$style(HTML("
+      .shiny-output-error-validation {
+        color: red;
+        font-size:120%;
+      },
+    ")),
+      # font for shiny notification
+      tags$style(
+        HTML(".shiny-notification {
[…]

### 2020-10-07 shiny-app/webinar-dec20/server.R
fixed violin & scatter bugs

@@ -704,11 +704,6 @@ server <- function(input, output,session) {
     ggplotly(plot)
     
     # ################################################
-    # ggplot(df) +
-    #  geom_mosaic(aes(x=product(gval1), fill=gval2),data=df)+xlab(v25$gvMosaic1)+ylab(v26$gvMosaic2)+ggtitle("Mosaic Plot")
-    # plot<-ggplot(df) +
-    #   geom_mosaic(aes(x=product(cots), fill=generation))+xlab(v25$gvMosaic1)+ylab(v26$gvMosaic2)
-    # ggplotly(plot)
     # something need to note is that ggplotly is not compatiable with geom_mosaic
   })
   
@@ -817,8 +812,9 @@ server <- function(input, output,session) {
     # use subset func to extract col object from dataframe; other func, such as df[...], seems not work at all
     # ggplot2 does not accept list object; must use unlist
     y_val<-unlist(subset(df, select=c(v30$qScatter)))
-    print(x_val)
-    p<-ggplot(df, aes(x = x_val, y = y_val, fill = x_val)) +
+    # x_val_cat=factor(x_val)
+    p<-ggplot(df, aes(x = factor(x_val), y = y_val, fill = factor(x_val))) +
+      xlab(v29$gvScatter)+labs(fill=v29$gvScatter)+ylab(v30$qScatter)+
       geom_point(pch = v34$pointShapeScatter, alpha=v33$transScatter/100, position = position_jitterdodge(jitter.height=0.075, jitter.width=0.1),size=v35$pointSizeScatter)+
       theme(
         plot.title = element_text(hjust=0.5, size=rel(1.8)),
@@ -827,32 +823,27 @@ server <- function(input, output,session) {
         axis.text.x = element_text(colour="grey", size=rel(1.5), angle=0, hjust=.5, vjust=.5, face="plain"),
         axis.text.y = element_text(colour="grey", size=rel(1.5), angle=0, hjust=.5, vjust=.5, face="plain"),
         panel.background = element_blank(),
-        axis.line = element_line(colour = "grey")##,
+        axis.line = element_line(colour = "grey"),
+        text=element_text(size=9)
       )
     
-    # p<-ggplot(df, aes(y = y_val, x = x_val, fill = x_val)) +
-    #   xlab(v29$gvScatter)+labs(fill=v29$gvScatter)+ylab(v30$qScatter)+
[…]

### 2020-10-07 shiny-app/webinar-dec20/ui.R

@@ -144,7 +144,7 @@ shinyUI(dashboardPage(
                              plotlyOutput("densities") ),
           ),
           box(
-            width = 4, status = "info",solidHeader = T,
+            width = 4, status = "info",solidHeader = T, 
             title = "Data Visualization",
             h4("Plot Option:"),
             

### 2020-10-08 ecosystem-data-analysis.Rmd

@@
 library(ggplot2)
-library(car)
 ```
@@
 ```{r, echo=FALSE}
-p2 = ggplot(dat2, aes(x = Osmocote, y = Height, fill = Osmocote)) +
+p2.2 = ggplot(dat2, aes(x = Osmocote, y = Height, fill = Osmocote)) +
   geom_point(pch = 21, alpha=0.3, position = position_jitterdodge(jitter.height=0.075, jitter.width=0.1))+
-  #ylim(c(1,6))+
+  geom_smooth(method='lm', formula= y~x)+
   theme(
@@
         )
-p2
+p2.2
 ```
+
+We can also do violin plots if one of the variables in categorical.
+We do not have any independent categorical variable right now (we only have osmocotes), but we could pretend that we have a categorical variable by treating the values of osmocotes as categories:
+```{r}
+dat2$osmocote.cat = factor(dat2$Osmocote)
+str(dat2)
+```
+
+```{r}
+p4 = ggplot(dat2, aes(x=osmocote.cat, y=Height, fill=osmocote.cat))+geom_violin(alpha=0.5)+
+  theme(
+        plot.title = element_text(hjust=0.5, size=rel(1.8)),
+        axis.title.x = element_text(size=rel(1.8)),
+        axis.title.y = element_text(size=rel(1.8), angle=90, vjust=0.5, hjust=0.5),
+        axis.text.x = element_text(colour="grey", size=rel(1.5), angle=0, hjust=.5, vjust=.5, face="plain"),
+        axis.text.y = element_text(colour="grey", size=rel(1.5), angle=0, hjust=.5, vjust=.5, face="plain"),
+        panel.background = element_blank(),
+        axis.line = element_line(colour = "grey")##,
+        )
+
+p4
+```

### 2020-11-04 shiny-app/webinar-aug20/ui.R
do some minor fix

@@ -64,7 +64,6 @@ shinyUI(dashboardPage(
                    
                    h4(HTML( 'WI Fast Stats is the <a href="https://github.com/crsl4/fast-stats" target="_blank">open-source</a> publicly available web app to analyze data from <a href="https://fastplants.org/" target="_blank">WI Fast Plants</a>.')),
                    h4(HTML( 'This web app is the accompanying tool for the WI Fast Plants webinar: <a href="https://fastplants.org/2020/08/06/new-fast-plants-polycots-selection/" target="_blank"><i>Strategies for adapting WI Fast Plants selection of traits investigations for remote and social distance learning</i></a>.')))
-              
       ),
       
       tabItem("dataUpload",

### 2020-11-04 shiny-app/webinar-aug20/ui.R
changed FAQ

@@ -374,10 +374,11 @@ shinyUI(dashboardPage(
                    h4("Q: How to get help? "), 
                    p(HTML('<b>A: Soon we will have a google user group to post questions and answers for users of the app.</b>')),
                    h4("Q: Webinar Links: "),  
-                   p(HTML('WI Fast Plants webinar: <a href="https://fastplants.org/2020/08/06/new-fast-plants-polycots-selection/" target="_blank"><i>Strategies for adapting WI Fast Plants selection of traits investigations for remote and social distance learning</i></a>.')),
+                   p(HTML('<b>A: WI Fast Plants webinar: <a href="https://fastplants.org/2020/08/06/new-fast-plants-polycots-selection/" target="_blank"><i>Strategies for adapting WI Fast Plants selection of traits investigations for remote and social distance learning</i></a>.</b>')),
                    h4("Q: Color Palettes Charts: "), 
+                   p(HTML('<b>A: The colors palettes here shown come from <a href="https://cran.r-project.org/web/packages/RColorBrewer/index.html">ColorBrewer</a></b>')),
                    img(src="color_palettes.png",width=525, height=671),
-                   p(HTML('   The colors palettes here shown come from <a href="https://cran.r-project.org/web/packages/RColorBrewer/index.html">ColorBrewer</a>')),
+                 
               )
       )
     ),#tabitems end

### 2020-11-04 shiny-app/webinar-dec20/ui.R

@@ -282,16 +282,15 @@ shinyUI(dashboardPage(
       ),
      
       tabItem("FAQ",
-              box( width=12, status="primary",solidHeader = T,
+              box( width=12, status="warning",solidHeader = T,
                    title="Frequently Asked Questions",
                    h4("Q: How to get help? "), 
                    p(HTML('<b>A: Soon we will have a google user group to post questions and answers for users of the app.</b>')),
                    h4("Q: Webinar Links: "),  
-                   p(HTML('WI Fast Plants webinar: <a href="https://fastplants.org/2020/08/06/new-fast-plants-polycots-selection/" target="_blank"><i>Strategies for adapting WI Fast Plants selection of traits investigations for remote and social distance learning</i></a>.')),
+                   p(HTML('<b>A: WI Fast Plants webinar: <a href="https://fastplants.org/2020/08/06/new-fast-plants-polycots-selection/" target="_blank"><i>Strategies for adapting WI Fast Plants selection of traits investigations for remote and social distance learning</i></a>.</b>')),
                    h4("Q: Color Palettes Charts: "), 
+                   p(HTML('<b>A: The colors palettes here shown come from <a href="https://cran.r-project.org/web/packages/RColorBrewer/index.html">ColorBrewer</a></b>')),
                    img(src="color_palettes.png",width=525, height=671),
-                   p(HTML('   The colors palettes here shown come from <a href="https://cran.r-project.org/web/packages/RColorBrewer/index.html">ColorBrewer</a>')),
-                  
                    
               )
       )

### 2020-11-04 cotyledon-data-analysis.Rmd
added mosaic plot

@@
 The p-value less than 0.05 allows us to also reject the null hypothesis of equality of distributions which provides evidence that the number of cotyledons on the offspring group is significantly higher than in the parent group.
+=======
+The p-value less than 0.05 allows us to also reject the null hypothesis of equality of distributions which provides evidence that the number of cotyledons on the offspring group is significantly higher than in the parent group.
 
@@
 ggplot(data = dat2) +
-  geom_mosaic(aes(x = product(cots), fill=generation))
+  geom_mosaic(aes(x=product(cots), fill=generation))
 ```

### 2020-11-04 shiny-app/webinar-aug20/ui.R
fixed minor issues

@@ -155,7 +155,7 @@ shinyUI(dashboardPage(
                              HTML('<p style="color:#808080"> <b>Mosaic plot:</b> Plot to visualize contigency tables of frequencies among categorical variables </p>'),
                              
                              h4("Data Visualization:"),
-                             HTML("Please select the two variables to use in the plot and click on the button to generate the plot."),
+                             HTML('<p style="color:#808080">Please select the two variables to use in the plot and click on the button to generate the plot.</p>'),
                              
                              fluidRow(
                                column(6,radioButtons("gvMosaic1","Group Variable 1:", c("1"="1","2"="2"))),
@@ -184,7 +184,7 @@ shinyUI(dashboardPage(
                              HTML('<p style="color:#808080"> <b>Add data points: </b> This option allows the user to add a scatterplot of the data where each dot corresponds to one observation (row) in the dataset. </p>'),
                              
                              h4("Data Visualization:"),
-                             HTML("Please select the two variables to use in the plot and click on the button to generate the plot."),
+                             HTML('<p style="color:#808080">Please select the two variables to use in the plot and click on the button to generate the plot.</p>'),
                              
                              fluidRow(
                                column(6,radioButtons("xaxisGrp","Group Variable:", c("1"="1","2"="2"))),
@@ -226,7 +226,7 @@ shinyUI(dashboardPage(
                              HTML('<p style="color:#808080"> <b>Add data points: </b>  This option allows the user to add a scatterplot of the data where each dot corresponds to one observation (row) in the dataset. </p>'),
                              
                              h4("Data Visualization:"),
-                             HTML("Please select the two variables to use in the plot and click on the button to generate the plot."),
+                             HTML('<p style="color:#808080">Please select the two variables to use in the plot and click on the button to generate the plot.</p>'),
                              
                              fluidRow(
                                column(6,radioButtons("gvBox","Group Variable:",c("1"="1","2"="2"))),
@@ -263,7 +263,7 @@ shinyUI(dashboardPage(
                              HTML('<p style="color:#808080"> <b>Scatter Plot:</b> Plot that shows the relationship between two numerical variables </p>'),
                              
                              h4("Data Visualization:"),
-                             HTML("Please select the two variables to use in the plot and click on the button to generate the plot."),
+                             HTML('<p style="color:#808080">Please select the two variables to use in the plot and click on the button to generate the plot.</p>'),
                              
                              fluidRow(
[…]

### 2020-11-04 shiny-app/webinar-dec20/ui.R

@@ -161,7 +161,7 @@ shinyUI(dashboardPage(
             conditionalPanel(condition="input.plotType==0",
                              
                              h4("Data Visualization:"),
-                             HTML("Please select the two variables to use in the plot and click on the button to generate the plot."),
+                             HTML('<p style="color:#808080">Please select the two variables to use in the plot and click on the button to generate the plot.</p>'),
                              
                              fluidRow(
                                column(6,radioButtons("gvMosaic1","Group Variable 1:", c("1"="1","2"="2"))),
@@ -191,9 +191,8 @@ shinyUI(dashboardPage(
                              HTML('<p style="color:#808080"> <b>Add data points: </b> This option allows the user to add a scatterplot of the data where each dot corresponds to one observation (row) in the dataset. </p>'),
                              
                              h4("Data Visualization:"),
-                             HTML("Please select the two variables to use in the plot and click on the button to generate the plot."),
-                             
                              HTML('<p style="color:#808080">Please select the two variables to use in the plot and click on the button to generate the plot.</p>'),
+                             
                              
                              fluidRow(
                                column(6,radioButtons("xaxisGrp","Group Variable:", c("1"="1","2"="2"))),
@@ -236,8 +235,6 @@ shinyUI(dashboardPage(
                              HTML('<p style="color:#808080"> <b>Add data points: </b>  This option allows the user to add a scatterplot of the data where each dot corresponds to one observation (row) in the dataset. </p>'),
                              
                              h4("Data Visualization:"),
-                             HTML("Please select the two variables to use in the plot and click on the button to generate the plot."),
-                             
                              HTML('<p style="color:#808080">Please select the two variables to use in the plot and click on the button to generate the plot.</p>'),
                              
                              fluidRow(
@@ -273,7 +270,6 @@ shinyUI(dashboardPage(
                              HTML('<p style="color:#808080"> <b>Scatter Plot:</b> Plot that shows the relationship between two numerical variables </p>'),
                              
                              h4("Data Visualization:"),
-                             HTML("Please select the two variables to use in the plot and click on the button to generate the plot."),
                              
                              HTML('<p style="color:#808080">Please select the two variables to use in the plot and click on the button to generate the plot.</p>'),
                              
@@ -317,7 +313,7 @@ shinyUI(dashboardPage(
       ),
      
       tabItem("FAQ",
-              box( width=12, status="warning",solidHeader = T,
[…]

### 2020-11-20 shiny-app/webinar-aug20/ui.R
minor changes to text in web apps

@@ -61,7 +61,7 @@ shinyUI(dashboardPage(
     tabItems(
       tabItem("about",
               box( width=12, status="warning",solidHeader = T,
-                   title = "Welcome to the WI Fast Stats app!",
+                   title = "Welcome to the WI Fast Stats app: Cotelydon!",
                    
                    h4(HTML( 'WI Fast Stats is the <a href="https://github.com/crsl4/fast-stats" target="_blank">open-source</a> publicly available web app to analyze data from <a href="https://fastplants.org/" target="_blank">WI Fast Plants</a>.')),
 
@@ -121,7 +121,7 @@ shinyUI(dashboardPage(
                   ),
                   
                   conditionalPanel(condition="input.fileType=='sampleFile'", 
-                                   HTML('<p style="color:#808080">The sample data included here mimics the structure of a dataset the students will have after following the experiments described in the webinar </p>'),
+                                   HTML('<p style="color:#808080">The sample data included here mimics the structure of a dataset the students will have after following the experiments described in the webinar. Each row corresponds to a plant, and we count the number of cotelydons. The generation column identifies if the plant is "parent" (p) or "offspring" (O). The parents column shows which were the parents of that plant. For example, 2x2 means that both parents had 2 cotelydons. </p>'),
                                    
                   ),
                   # Input: Select number of rows to display ----
@@ -383,7 +383,7 @@ shinyUI(dashboardPage(
                              HTML('<p style="color:#808080"> <b>Chi-square test:</b> Pearson\'s chi-square test is used to determine whether there is a statistically significant difference between the expected frequencies and the observed frequencies in one or more categories of a contingency table</p>'),
                              fluidRow(
                                column(6, align="center", offset = 3,
-                                      actionButton("goChi", "Chi Sqaure Test"),
+                                      actionButton("goChi", "Chi Square Test"),
                                       tags$style(type='text/css', "#button { vertical-align: middle; height: 50px; width: 100%; font-size: 30px;}")
                                )
                              ),
@@ -397,8 +397,8 @@ shinyUI(dashboardPage(
               box( width=12, status="warning",solidHeader = T,
                    title="Frequently Asked Questions",
                    h4("Q: How to get help? "), 
-                   p(HTML('<b>A: Soon we will have a google user group to post questions and answers for users of the app.</b>')),
[…]

### 2020-11-20 shiny-app/webinar-dec20/ui.R

@@ -64,11 +64,11 @@ shinyUI(dashboardPage(
     tabItems(
       tabItem("about",
               box( width=12, status="primary",solidHeader = T,
-                   title = "Welcome to the WI Fast Stats app!",
+                   title = "Welcome to the WI Fast Stats app: Ecosystem!",
                    
                    h4(HTML( 'WI Fast Stats is the <a href="https://github.com/crsl4/fast-stats" target="_blank">open-source</a> publicly available web app to analyze data from <a href="https://fastplants.org/" target="_blank">WI Fast Plants</a>.')),
 
-                   h4(HTML( 'This web app is the accompanying tool for the WI Fast Plants webinar: <a href="https://fastplants.org/2020/08/06/new-fast-plants-polycots-selection/" target="_blank"><i>Strategies for adapting WI Fast Plants selection of traits investigations for remote and social distance learning</i></a>.')))
+                   h4(HTML( 'This web app is the accompanying tool for the WI Fast Plants webinar: <a href="https://register.gotowebinar.com/register/1900289654623158541" target="_blank"><i>Teaching Middle-level and Upper-elementary Students Ecosystem Concepts with Hands-on Fast Plants Investigations</i></a>.')))
 
               
       ),
@@ -317,9 +317,9 @@ shinyUI(dashboardPage(
               box( width=12, status="primary",solidHeader = T,
                    title="Frequently Asked Questions",
                    h4("Q: How to get help? "), 
-                   p(HTML('<b>A: Soon we will have a google user group to post questions and answers for users of the app.</b>')),
-                   h4("Q: Webinar Links: "),  
-                   p(HTML('<b>A: WI Fast Plants webinar: <a href="https://fastplants.org/2020/08/06/new-fast-plants-polycots-selection/" target="_blank"><i>Strategies for adapting WI Fast Plants selection of traits investigations for remote and social distance learning</i></a>.</b>')),
+                   p(HTML('<b>A: Check out the WI Fast Stats google user group where people post questions/answers. You can join to post questions: <a href="https://groups.google.com/g/wi-fast-stats/">https://groups.google.com/g/wi-fast-stats</a></b>')),
+                   h4("Q: Where can I find the information about the WI Fast Plants Webinar?"),  
[…]

### 2020-11-20 notebooks/ecosystem-data-analysis.Rmd

--- 2020-10-08 ecosystem-data-analysis.Rmd
@@
   geom_point(pch = 21, alpha=0.3, position = position_jitterdodge(jitter.height=0.075, jitter.width=0.1))+
+  #scale_fill_gradient2()+
+  scale_colour_gradientn(colours = terrain.colors(10)) +
   theme(
@@
 p2
+
+## testing the scale_fill_gradient
+library(RColorBrewer)
+group_list=unlist(subset(dat2,select=c(as.factor(Osmocote))))
+colorCount = length(unique(unlist(group_list,use.names=F)))
+getPalette <- colorRampPalette(brewer.pal(8, "OrRd"),bias=2.5)(colorCount)
+
+p2 + scale_fill_gradient(low=getPalette[1], high=getPalette[length(getPalette)])
 ```

### 2020-11-21 shiny-app/webinar-dec20/ui.R
some minor changes

@@ -117,6 +117,9 @@ shinyUI(dashboardPage(
                   
                   conditionalPanel(condition="input.fileType=='sampleFile'", 
                                    HTML('<p style="color:#808080">The sample data included here mimics the structure of a dataset the students will have after following the experiments described in the webinar </p>'),
+                                   HTML('<p style="color:#808080">Each row corresponds to a plant. The treatment column refers to dark (D) or light (L) and we measure different lengths: root, shoot and total</p>'),
+                                   
+                                   
                                    
                   ),
                   # Input: Select number of rows to display ----
@@ -187,10 +190,6 @@ shinyUI(dashboardPage(
                              
                              HTML('<p style="color:#808080"> <b>Violin Plot:</b> Plot a numerical variable ("Quantity") by groups ("Group variable"). It is similar to the box plot but it also shows the distribution and spread of the data. </p>'),
                              
-                             checkboxInput("addPoints", "Add data points", T),
-                             
-                             HTML('<p style="color:#808080"> <b>Add data points: </b> This option allows the user to add a scatterplot of the data where each dot corresponds to one observation (row) in the dataset. </p>'),
-                             
                              h4("Data Visualization:"),
                              HTML('<p style="color:#808080">Please select the two variables to use in the plot and click on the button to generate the plot.</p>'),
                              
@@ -200,11 +199,16 @@ shinyUI(dashboardPage(
                                column(6,radioButtons("yaxisGrp","Quantity:", c("1"="1","2"="2")))
                              ),
                              
+                             checkboxInput("addPoints", "Add data points", T),
+                             
+                             HTML('<p style="color:#808080"> <b>Add data points: </b> This option allows the user to add a scatterplot of the data where each dot corresponds to one observation (row) in the dataset. </p>'),
+                             
                              selectInput("colorViolin","Color",choices = c("Blue+Purple"="BuPu","Dark Color"="Dark2","Orange+Red"="OrRd","Yellow+Green+Blue"="YlGnBu","Accent"="Accent","Paired"="Paired","Red+Blue"="RdYlBu","Purple+Red"="PuRd","Set2"="Set2","Purple+Green"="PRGn")
                              ),
 
                              HTML('<p style="color:#808080"> <b>About colors:</b> Click the <b>FAQ</b> tab in the left sidebar for more details on the color palettes</p>'),
[…]

### 2020-11-22 shiny-app/webinar-aug20/ui.R
add FAQ questions

@@ -408,6 +408,8 @@ shinyUI(dashboardPage(
                    p(HTML('<b>A: WI Fast Plants webinar: <a href="https://fastplants.org/2020/08/06/new-fast-plants-polycots-selection/" target="_blank"><i>Strategies for adapting WI Fast Plants selection of traits investigations for remote and social distance learning</i></a>.</b>')),
                    h4("Q: Where can I find the webinar slides for the Data science part?"),
                    p(HTML('<b>A: The webinar slides are in the WI Fast Stats github repo <a href="https://github.com/crsl4/fast-stats" target="_blank">here</a></b>')),
+                   h4("Q: I found a bug or error in the code, how can I report it?"),
+                   p(HTML('<b>A: You should file an issue in the github repo: <a href="https://github.com/crsl4/fast-stats/issues" target="_blank">https://github.com/crsl4/fast-stats/issues</a></b>')),
                    h4("Q: How can I provide positive (or constructive) feedback?"),
                    p(HTML(' <b>A: Users feedback is very important to us! Please use <a href="https://forms.gle/PQS92afdhLHwK6kG6" target="_blank">this form </a></b>')),
                    h4("Q: If I use the website and web apps in my work, how do I cite them?"),

### 2022-03-19 ms/response2reviews.md
modified license on code comments

We thank the reviewers for such helpful comments that have greatly improved the web apps and the manuscript.

# Review 1

Functionality and Documentation:
- We double checked all the steps in the `readme` file to run the web apps locally, and provided more details on how to install `thatssorandom` and how to run the `shiny::runApp` command
- We have added more comments to the code and we have added a new section on the documentation for young audiences interested in coding
- We modified the interface so that only categorical variables appear under "Group variable" and only quantitative variables appear under "quantity"
- For the `webinar-aug20` sample data, we make `plant ID` a group variable. We found that for the other web apps, it did not make sense to change `plant ID` as categorical variable because some plots became not very intepretable, so we decided to keep them as quantitiative variables.

add more details on how to install thatssorandom
add more details on how to run runApp
overall check all the steps to run a web app locally
more comments on code! Think of younger audiences looking at the code. No need to comment every line, but explain the overall structure and what each function is doing
“Group variable” should only display the variables that are categorical and “Quantity” should only display the variables that are quantitative
For cotelydon toy data, change plant ID to categorical variable
change the warning on guide scales=FALSE to “none”
check the warnings on fewer than two data points

Add DOI to references
Change Data summary to Data upload
Expand documentation to help students
Make slides available and data for the slides: Maybe change to md
remove the notebooks
editorial comments on broken links

### 2022-03-19 shiny-app/webinar-aug20/server.R

@@ -5,16 +5,9 @@
 # 
 #
 #     This program is free software; you can redistribute it and/or
-#     modify it under the terms of the GNU General Public License,
-#     version 3, as published by the Free Software Foundation.
+#     modify it under the terms of the MIT License:
+#     https://github.com/crsl4/fast-stats/blob/master/LICENSE
 #
-#     This program is distributed in the hope that it will be useful,
-#     but without any warranty; without even the implied warranty of
-#     merchantability or fitness for a particular purpose.  See the GNU
-#     General Public License, version 3, for more details.
-#
-#     A copy of the GNU General Public License, version 3, is available
-#     at http://www.r-project.org/Licenses/GPL-3
 #
 # Part of the fast-stats package
 # Contains: server.R
@@ -39,7 +32,7 @@ toy<-read.csv("toys.csv",
               header = T,
               sep = ",")
 ######################################################################
-# server: R shiny server function, 
+# server: R shiny server function (see tutorial: https://shiny.rstudio.com/tutorial/)
 # 
 # input:
 # The session's input object (the same as is passed into the Shiny server 

### 2022-03-19 shiny-app/webinar-aug20/ui.R

@@ -5,16 +5,8 @@
 # 
 #
 #     This program is free software; you can redistribute it and/or
-#     modify it under the terms of the GNU General Public License,
-#     version 3, as published by the Free Software Foundation.
-#
-#     This program is distributed in the hope that it will be useful,
-#     but without any warranty; without even the implied warranty of
-#     merchantability or fitness for a particular purpose.  See the GNU
-#     General Public License, version 3, for more details.
-#
-#     A copy of the GNU General Public License, version 3, is available
-#     at http://www.r-project.org/Licenses/GPL-3
+#     modify it under the terms of the MIT License:
+#     https://github.com/crsl4/fast-stats/blob/master/LICENSE
 #
 # Part of the fast-stats package
 # Contains: ui.R
@@ -26,9 +18,9 @@ library(plotly)
 library(shinydashboard)
 ######################################################################
 # 
-# shinyUI: Create a Shiny UI handler
-# ui:
-# A user interace definition
+# shinyUI: Create a Shiny UI handler (see tutorial: https://shiny.rstudio.com/tutorial/)
+# ui: user interface. This function will define the format of the 
+# screen (what the user sees when opening the web app)
 # 
 ###################################################################### 
 shinyUI(dashboardPage(

### 2022-03-19 shiny-app/webinar-bio152/server.R

@@ -5,16 +5,8 @@
 # 
 #
 #     This program is free software; you can redistribute it and/or
-#     modify it under the terms of the GNU General Public License,
-#     version 3, as published by the Free Software Foundation.
-#
-#     This program is distributed in the hope that it will be useful,
-#     but without any warranty; without even the implied warranty of
-#     merchantability or fitness for a particular purpose.  See the GNU
-#     General Public License, version 3, for more details.
-#
-#     A copy of the GNU General Public License, version 3, is available
-#     at http://www.r-project.org/Licenses/GPL-3
+#     modify it under the terms of the MIT License:
+#     https://github.com/crsl4/fast-stats/blob/master/LICENSE
 #
 # Part of the fast-stats package
 # Contains: server.R
@@ -38,7 +30,7 @@ toy<-read.csv("2016-brassica-new.csv",
               header = T,
               sep = ",")
 ######################################################################
-# server: R shiny server function, 
+# server: R shiny server function (see tutorial: https://shiny.rstudio.com/tutorial/)
 # 
 # input:
 # The session's input object (the same as is passed into the Shiny server 

### 2022-03-19 shiny-app/webinar-bio152/ui.R

@@ -5,16 +5,8 @@
 # 
 #
 #     This program is free software; you can redistribute it and/or
-#     modify it under the terms of the GNU General Public License,
-#     version 3, as published by the Free Software Foundation.
-#
-#     This program is distributed in the hope that it will be useful,
-#     but without any warranty; without even the implied warranty of
-#     merchantability or fitness for a particular purpose.  See the GNU
-#     General Public License, version 3, for more details.
-#
-#     A copy of the GNU General Public License, version 3, is available
-#     at http://www.r-project.org/Licenses/GPL-3
+#     modify it under the terms of the MIT License:
+#     https://github.com/crsl4/fast-stats/blob/master/LICENSE
 #
 # Part of the fast-stats package
 # Contains: ui.R
@@ -27,9 +19,9 @@ library(shinydashboard)
 
 ######################################################################
 # 
-# shinyUI: Create a Shiny UI handler
-# ui:
-# A user interace definition
+# shinyUI: Create a Shiny UI handler (see tutorial: https://shiny.rstudio.com/tutorial/)
+# ui: user interface. This function will define the format of the 
+# screen (what the user sees when opening the web app)
 # 
 ###################################################################### 
 shinyUI(dashboardPage(

### 2022-03-19 shiny-app/webinar-dec20/server.R

@@ -5,16 +5,8 @@
 # 
 #
 #     This program is free software; you can redistribute it and/or
-#     modify it under the terms of the GNU General Public License,
-#     version 3, as published by the Free Software Foundation.
-#
-#     This program is distributed in the hope that it will be useful,
-#     but without any warranty; without even the implied warranty of
-#     merchantability or fitness for a particular purpose.  See the GNU
-#     General Public License, version 3, for more details.
-#
-#     A copy of the GNU General Public License, version 3, is available
-#     at http://www.r-project.org/Licenses/GPL-3
+#     modify it under the terms of the MIT License:
+#     https://github.com/crsl4/fast-stats/blob/master/LICENSE
 #
 # Part of the fast-stats package
 # Contains: server.R
@@ -36,7 +28,7 @@ toy<-read.csv("2016-brassica.csv",
              header = T,
              sep = ",")
 ######################################################################
-# server: R shiny server function, 
+# server: R shiny server function  (see tutorial: https://shiny.rstudio.com/tutorial/)
 # 
 # input:
 # The session's input object (the same as is passed into the Shiny server 

### 2022-03-19 shiny-app/webinar-dec20/ui.R

@@ -5,16 +5,8 @@
 # 
 #
 #     This program is free software; you can redistribute it and/or
-#     modify it under the terms of the GNU General Public License,
-#     version 3, as published by the Free Software Foundation.
-#
-#     This program is distributed in the hope that it will be useful,
-#     but without any warranty; without even the implied warranty of
-#     merchantability or fitness for a particular purpose.  See the GNU
-#     General Public License, version 3, for more details.
-#
-#     A copy of the GNU General Public License, version 3, is available
-#     at http://www.r-project.org/Licenses/GPL-3
+#     modify it under the terms of the MIT License:
+#     https://github.com/crsl4/fast-stats/blob/master/LICENSE
 #
 # Part of the fast-stats package
 # Contains: ui.R
@@ -27,9 +19,9 @@ library(shinyjs)
 
 ######################################################################
 # 
-# shinyUI: Create a Shiny UI handler
-# ui:
-# A user interace definition
+# shinyUI: Create a Shiny UI handler (see tutorial: https://shiny.rstudio.com/tutorial/)
+# ui: user interface. This function will define the format of the 
+# screen (what the user sees when opening the web app)
 # 
 ###################################################################### 
 shinyUI(dashboardPage(

### 2022-03-19 ms/response2reviews.md
changes to manuscript and references

@@
 
-add more details on how to install thatssorandom
-add more details on how to run runApp
-overall check all the steps to run a web app locally
-more comments on code! Think of younger audiences looking at the code. No need to comment every line, but explain the overall structure and what each function is doing
-“Group variable” should only display the variables that are categorical and “Quantity” should only display the variables that are quantitative
-For cotelydon toy data, change plant ID to categorical variable
-change the warning on guide scales=FALSE to “none”
-check the warnings on fewer than two data points
+Software paper
+- We added the DOI to me Williams and Hill (1986) paper
+- We now make reference to the "Data Upload" section in the paper instead of the "Data Summary"
 
-Add DOI to references
-Change Data summary to Data upload
+# Review 2
+
+Functionality and Documentation:
+- We double checked all the steps in the `readme` file to run the web apps locally, and provided more details on how to install `thatssorandom` and how to run the `shiny::runApp` command
+- Unfortunately, the `Warning: guides(<scale> = FALSE) is deprecated.
+Please use guides(<scale> = "none") instead.` is out of our control as we are not calling `guides` directly in our mosaic function.
+We wondered whether we should suppress all warnings so that the user is not confused, but decided against this in case there are bugs or package upgrades that we need to address in the future.
+For now, we simply added a section about warnings in the documentation.
+- We modified the interface so that only categorical variables appear under "Group variable" and only quantitative variables appear under "quantity"
+- We have also expanded the documentation including more details about the sample dataset and some of the most common errors and warnings
+- The slides provided in the repo are the slides used when teaching the WI Fast plants webinars.
+The plots in the slides were not created using the same sample data in the repo (but data related to the specific webinar).
+The goal of the slides was to illustrate concepts, rather than testing the web app.
+This is actually related to another point brought up by the reviewer (the notebooks).
+It was precisely the goal of the notebooks to be that reproducible script that anyone could run (but as the reviewer mentioned, this was not true).
+In short, the slides are made available as part of the webinars to illustrate statistical concepts rather than usage of the web apps.
+I have now updated the notebooks so that they are indeed reproducible so that anyone can run them and get the plots.
+Hope this satisfies the reviewer!
+
+Notebooks folder
+- We have double checked the steps to the notebook to make sure they are indeed reproducible and we have included a readme file explaining how the notebooks can be used to reproduce plots and analyses
+
+Software paper
+- We have added the DOI to all papers
+
+Editorial
+- We fully revised the documentation
+- Thanks for checking that the links to the webinars do not work anymore.
+The webinars were run in coordination with WI Fast Plants and Carolina Knowledge Center, so the links are on their end.
+We have reached out to them to ask whether the links have changed, but we have not heard back yet.
+
 Expand documentation to help students
-Make slides available and data for the slides: Maybe change to md
-remove the notebooks
-editorial comments on broken links
+  - section on younger audiences starting to code: point at shiny app tutorial
+  - section on warnings in the code: Warning: Groups with fewer than two data points have been dropped
+  - section on sample data
+  - section on most common errors
+make the notebooks reproducible (change to sample data)
+  - distinguish which are for webinar and which are for reproducing
+  - add readme to slides that the slides are meant to illustrate stat concepts

### 2022-03-19 ms/response2reviews.md
changes to docs

@@
 The plots in the slides were not created using the same sample data in the repo (but data related to the specific webinar).
-The goal of the slides was to illustrate concepts, rather than testing the web app.
+The goal of the slides was to illustrate statistical concepts, rather than testing the web app.
 This is actually related to another point brought up by the reviewer (the notebooks).
-It was precisely the goal of the notebooks to be that reproducible script that anyone could run (but as the reviewer mentioned, this was not true).
-In short, the slides are made available as part of the webinars to illustrate statistical concepts rather than usage of the web apps.
-I have now updated the notebooks so that they are indeed reproducible so that anyone can run them and get the plots.
+It was precisely the goal of the notebooks to be that reproducible script that anyone could run (but as the reviewer mentioned, they were not working properly).
+We have now updated the notebooks so that they are indeed reproducible so that anyone can run them and get the plots.
+We have also added a readme file explaining this in the repo.
 Hope this satisfies the reviewer!
@@
 We have reached out to them to ask whether the links have changed, but we have not heard back yet.
-
-Expand documentation to help students
-  - section on younger audiences starting to code: point at shiny app tutorial
-  - section on warnings in the code: Warning: Groups with fewer than two data points have been dropped
-  - section on sample data
-  - section on most common errors
-make the notebooks reproducible (change to sample data)
-  - distinguish which are for webinar and which are for reproducing
-  - add readme to slides that the slides are meant to illustrate stat concepts
+If it is ok with the reviewers, we will create a new issue on the broken links for now to address this when we receive the new links.

### 2022-03-19 notebooks/cotyledon-data-analysis.Rmd
changes to notebooks

--- 2020-07-31 cotyledon-data-analysis.Rmd
@@
 
-# Initial data: `Polycot-data-from-NY.csv`
+# Data: `Polycot-data-2020.csv` focusing on chi-square test (webinar)
 
@@
 ```{r}
-dat = read.table("data/Polycot-data-from-NY.csv", sep=",",header=TRUE)
+dat = read.table("../data/Polycot-data-2020.csv", sep=",",header=TRUE)
+str(dat)
+head(dat)
+summary(dat)
+dat = within(dat, Generation<-factor(Generation, levels=c("P","O")))
+```
+
+Creating categorical variable: dycots/polycots
+
+```{r}
+dat2 = data.frame(generation=dat$Generation)
+dat2$cots = "dicot"
+dat2$cots[dat$Cotyledons > 2] = "polycot"
+dat2 = within(dat2, cots<-factor(cots))
+summary(dat2)
+#write.table(dat2,file="data/polycot-data-2020-categories.csv", row.names = FALSE, sep=",")
+
+dt = table(dat2$cots, dat2$generation)
+```
+
+### Data visualization
+```{r}
+library(graphics)
+mosaicplot(dt, shade = TRUE, las=2, main="Cotelydons")
+
+library(ggmosaic)
+ggplot(data = dat2) +
+  geom_mosaic(aes(x=product(cots), fill=generation))
+```
+
+### Data analysis
+```{r}
+dt
+ct = chisq.test(dt)
+
+obs = ct$observed
+exp = ct$expected
+
+df = rbind(obs[1,],exp[1,],obs[2,],exp[2,],colSums(obs))
+rownames(df) = c("dycot observed", "dycot expected", "polycot observed", "polycot expected", "total")
+
+r = rowSums(obs)
+ss = sum(r)
+totals = c(r[1],NA,r[2],NA,ss)
+cbind(df,totals)
+ct$statistic
+ct$p.value
+
+## expected values
+2437*2556/5585
+2437*3029/5585
+3148*2556/5585
+3148*3029/5585
+```
+
+# Data: `Polycot-data-2020.csv` (topics not covered in the webinar)
+
+The code of this section contains plots and statistical tests that were not included in the final version of the webinar.
+For the webinar, we decided to focus on the chi-square test.
+Here, we fit a t test, a Levene test (for equality of variances) and a Wilcoxon-Mann-Whitney test which is similar to the t test when normality is violated.
+We also fit an ANOVA test and its alternative when normality is not met, the Kruskal-Wallis rank sum test.
+
+### Reading the data
+
+Reading in the data (had to remove manually the '#' sign from the column and save as csv file):
+```{r}
+dat = read.table("../data/Polycot-data-2020.csv", sep=",",header=TRUE)
+str(dat)
+head(dat)
+summary(dat)
+dat = within(dat, Generation<-factor(Generation, levels=c("P","O")))
+```
+
+### Data visualization
+
+```{r, echo=FALSE}
+p1 = ggplot(dat, aes(Generation, fill=Generation))+geom_bar()+
+  theme(
+        plot.title = element_text(hjust=0.5, size=rel(1.8)),
+        axis.title.x = element_text(size=rel(1.8)),
+        axis.title.y = element_text(size=rel(1.8), angle=90, vjust=0.5, hjust=0.5),
+        axis.text.x = element_text(colour="grey", size=rel(1.5), angle=0, hjust=.5, vjust=.5, face="plain"),
+        axis.text.y = element_text(colour="grey", size=rel(1.5), angle=0, hjust=.5, vjust=.5, face="plain"),
+        panel.background = element_blank(),
+        axis.line = element_line(colour = "grey")##,
+        )
+p1
+```
+
+```{r, echo=FALSE}
+p3 = ggplot(dat, aes(x = Generation, y = Cotyledons, fill = Generation)) +
+  geom_jitter(pch = 21, alpha=0.3, height=0.2)+
+  theme(
+        plot.title = element_text(hjust=0.5, size=rel(1.8)),
+        axis.title.x = element_text(size=rel(1.8)),
+        axis.title.y = element_text(size=rel(1.8), angle=90, vjust=0.5, hjust=0.5),
+        axis.text.x = element_text(colour="grey", size=rel(1.5), angle=0, hjust=.5, vjust=.5, face="plain"),
+        axis.text.y = element_text(colour="grey", size=rel(1.5), angle=0, hjust=.5, vjust=.5, face="plain"),
+        panel.background = element_blank(),
+        axis.line = element_line(colour = "grey")##,
+        )
+p3
+```
+
+```{r, echo=FALSE}
+p3.2 = ggplot(dat, aes(x = Generation, y = Cotyledons, fill = Generation)) +
+  geom_jitter(pch = 21, alpha=0.3, height=0.2)+
+    geom_boxplot(outlier.size = 0, alpha=0.1) +
+  theme(
+        plot.title = element_text(hjust=0.5, size=rel(1.8)),
+        axis.title.x = element_text(size=rel(1.8)),
+        axis.title.y = element_text(size=rel(1.8), angle=90, vjust=0.5, hjust=0.5),
+        axis.text.x = element_text(colour="grey", size=rel(1.5), angle=0, hjust=.5, vjust=.5, face="plain"),
+        axis.text.y = element_text(colour="grey", size=rel(1.5), angle=0, hjust=.5, vjust=.5, face="plain"),
+        panel.background = element_blank(),
+        axis.line = element_line(colour = "grey")##,
+        )
+p3.2
+```
+
+```{r, echo=FALSE}
+p4 = ggplot(dat, aes(x=Generation, y=Cotyledons, fill=Generation))+geom_violin(alpha=0.5)+
+  ylim(c(1,6))+
+  theme(
+        plot.title = element_text(hjust=0.5, size=rel(1.8)),
+        axis.title.x = element_text(size=rel(1.8)),
+        axis.title.y = element_text(size=rel(1.8), angle=90, vjust=0.5, hjust=0.5),
+        axis.text.x = element_text(colour="grey", size=rel(1.5), angle=0, hjust=.5, vjust=.5, face="plain"),
+        axis.text.y = element_text(colour="grey", size=rel(1.5), angle=0, hjust=.5, vjust=.5, face="plain"),
+        panel.background = element_blank(),
+        axis.line = element_line(colour = "grey")##,
+        )
+
+p4
+```
+
+```{r, echo=FALSE}
+p4.2 = ggplot(dat, aes(x=Generation, y=Cotyledons, fill=Generation))+geom_violin(alpha=1)+
+    geom_jitter(pch = 21, alpha=0.1)+
+  ylim(c(1,6))+
+  theme(
+        plot.title = element_text(hjust=0.5, size=rel(1.8)),
+        axis.title.x = element_text(size=rel(1.8)),
+        axis.title.y = element_text(size=rel(1.8), angle=90, vjust=0.5, hjust=0.5),
+        axis.text.x = element_text(colour="grey", size=rel(1.5), angle=0, hjust=.5, vjust=.5, face="plain"),
+        axis.text.y = element_text(colour="grey", size=rel(1.5), angle=0, hjust=.5, vjust=.5, face="plain"),
+        panel.background = element_blank(),
+        axis.line = element_line(colour = "grey")##,
+        )
+
+p4.2
+```
+
+## Statistical tests
+
+### Standard t-test
+
+Standard t-test comparing:
+
+- Population 1: Parents
+- Population 2: Offspring
+
+Assumptions:
+
+- Data distributed normal
+- Equal variance in the two groups (we can test this with the Levene's test later)
+
+Plotting the densities:
+
+```{r, echo=FALSE}
+q1 = ggplot(dat, aes(Cotyledons, fill=Generation))+geom_density(alpha=0.25)+
+  theme(
+        plot.title = element_text(hjust=0.5, size=rel(1.8)),
+        axis.title.x = element_text(size=rel(1.8)),
+        axis.title.y = element_text(size=rel(1.8), angle=90, vjust=0.5, hjust=0.5),
+        axis.text.x = element_text(colour="grey", size=rel(1.5), angle=0, hjust=.5, vjust=.5, face="plain"),
+        axis.text.y = element_text(colour="grey", size=rel(1.5), angle=0, hjust=.5, vjust=.5, face="plain"),
+        panel.background = element_blank(),
+        axis.line = element_line(colour = "grey")##,
+        )
+q1
+```
+
+Not really normal!
+
+```{r}
+x = dat$Cotyledons[dat$Generation == "P"]
+y = dat$Cotyledons[dat$Generation == "O"]
+```
+
+Variances look similar:
+```{r}
+var(x)
+var(y)
+```
+
+But let's do a Levene test:
+```{r}
+leveneTest(Cotyledons ~ Generation, dat, center=mean)
+```
+
+The p-value is less than 0.05, so we reject the null hypothesis of equal variances.
+
+The Levene test also assumes Normality. So, we can run the alternative Fligner-Killeen test with the same conclusion (equal variances):
+```{r}
+fligner.test(Cotyledons ~ Generation, dat)
+```
+
+Back to the t-test assumptions:
+
+- Normality: not met
+- Equality of variances: not met
+
+If we ignore the violations and we run a standard t-test, we get:
+```{r}
+t.test(x,y, var.equal=TRUE)
+```
+
+The p-value is less than 0.05, so we reject the null hypothesis of equality of means.
+This implies that we have evidence that the number of cotyledons is significantly larger in the offspring than in the parents.
+
+Our justification to do a standard t-test is that if the sample size is moderately large, the two-sample t-test is robust to non-normality due to the central limit theorem.
+
+We can also run the Welch version of t test that does not assume equal variances (this is the default in R):
+
+```{r}
+t.test(x,y)
+```
+
+### Wilcoxon-Mann-Whitney test
+
+Alternative to the t-test when normality is violated.
+This test does not test equality of means, but equality of distributions, so the conclusion could be deceiving (that is, two different distributions with the same mean will have a rejected null hypothesis using WMW test).
+
+```{r}
+wilcox.test(Cotyledons ~ Generation, dat)
+```
+
+The p-value less than 0.05 allows us to also reject the null hypothesis of equality of distributions which provides evidence that the number of cotyledons on the offspring group is significantly higher than in the parent group.
+
+# Initial data `Polycot-data-from-NY.csv` (not publicly available)
+
+The following sections have code that was written before the final topics of the webinar were decided.
+The code below is only kept for internal book-keeping as the data is not made publicly available (this is not the data used in the webinar).
+Unlike the main section above, here we do not fit a chi-square test, but a t test, a Levene test (for equality of variances) and a Wilcoxon-Mann-Whitney test which is similar to the t test when normality is violated.
+We also fit an ANOVA test and its alternative when normality is not met, the Kruskal-Wallis rank sum test.
+
+### Reading the data
+
+Reading in the data (had to remove manually the '#' sign from the column and save as csv file):
+```{r}
+dat = read.table("../data/Polycot-data-from-NY.csv", sep=",",header=TRUE)
 str(dat)
@@
 ```
-
-# New data: `Polycot-data-2020.csv`
-
-### Reading the data
-
-Reading in the data (had to remove manually the '#' sign from the column and save as csv file):
-```{r}
-dat = read.table("data/Polycot-data-2020.csv", sep=",",header=TRUE)
-str(dat)
-head(dat)
-summary(dat)
-dat = within(dat, Generation<-factor(Generation, levels=c("P","O")))
-```
-
-### Data visualization
-
-```{r, echo=FALSE}
-p1 = ggplot(dat, aes(Generation, fill=Generation))+geom_bar()+
-  theme(
-        plot.title = element_text(hjust=0.5, size=rel(1.8)),
-        axis.title.x = element_text(size=rel(1.8)),
-        axis.title.y = element_text(size=rel(1.8), angle=90, vjust=0.5, hjust=0.5),
-        axis.text.x = element_text(colour="grey", size=rel(1.5), angle=0, hjust=.5, vjust=.5, face="plain"),
-        axis.text.y = element_text(colour="grey", size=rel(1.5), angle=0, hjust=.5, vjust=.5, face="plain"),
-        panel.background = element_blank(),
-        axis.line = element_line(colour = "grey")##,
-        )
-p1
-```
-
-```{r, echo=FALSE}
-p3 = ggplot(dat, aes(x = Generation, y = Cotyledons, fill = Generation)) +
-  geom_jitter(pch = 21, alpha=0.3, height=0.2)+
-  theme(
-        plot.title = element_text(hjust=0.5, size=rel(1.8)),
-        axis.title.x = element_text(size=rel(1.8)),
-        axis.title.y = element_text(size=rel(1.8), angle=90, vjust=0.5, hjust=0.5),
-        axis.text.x = element_text(colour="grey", size=rel(1.5), angle=0, hjust=.5, vjust=.5, face="plain"),
-        axis.text.y = element_text(colour="grey", size=rel(1.5), angle=0, hjust=.5, vjust=.5, face="plain"),
-        panel.background = element_blank(),
-        axis.line = element_line(colour = "grey")##,
-        )
-p3
-```
-
-```{r, echo=FALSE}
-p3.2 = ggplot(dat, aes(x = Generation, y = Cotyledons, fill = Generation)) +
-  geom_jitter(pch = 21, alpha=0.3, height=0.2)+
-    geom_boxplot(outlier.size = 0, alpha=0.1) +
-  theme(
-        plot.title = element_text(hjust=0.5, size=rel(1.8)),
-        axis.title.x = element_text(size=rel(1.8)),
-        axis.title.y = element_text(size=rel(1.8), angle=90, vjust=0.5, hjust=0.5),
-        axis.text.x = element_text(colour="grey", size=rel(1.5), angle=0, hjust=.5, vjust=.5, face="plain"),
-        axis.text.y = element_text(colour="grey", size=rel(1.5), angle=0, hjust=.5, vjust=.5, face="plain"),
-        panel.background = element_blank(),
-        axis.line = element_line(colour = "grey")##,
-        )
-p3.2
-```
-
-```{r, echo=FALSE}
-p4 = ggplot(dat, aes(x=Generation, y=Cotyledons, fill=Generation))+geom_violin(alpha=0.5)+
-  ylim(c(1,6))+
-  theme(
-        plot.title = element_text(hjust=0.5, size=rel(1.8)),
-        axis.title.x = element_text(size=rel(1.8)),
-        axis.title.y = element_text(size=rel(1.8), angle=90, vjust=0.5, hjust=0.5),
-        axis.text.x = element_text(colour="grey", size=rel(1.5), angle=0, hjust=.5, vjust=.5, face="plain"),
-        axis.text.y = element_text(colour="grey", size=rel(1.5), angle=0, hjust=.5, vjust=.5, face="plain"),
-        panel.background = element_blank(),
-        axis.line = element_line(colour = "grey")##,
-        )
-
-p4
-```
-
-```{r, echo=FALSE}
-p4.2 = ggplot(dat, aes(x=Generation, y=Cotyledons, fill=Generation))+geom_violin(alpha=1)+
-    geom_jitter(pch = 21, alpha=0.1)+
-  ylim(c(1,6))+
-  theme(
-        plot.title = element_text(hjust=0.5, size=rel(1.8)),
-        axis.title.x = element_text(size=rel(1.8)),
-        axis.title.y = element_text(size=rel(1.8), angle=90, vjust=0.5, hjust=0.5),
-        axis.text.x = element_text(colour="grey", size=rel(1.5), angle=0, hjust=.5, vjust=.5, face="plain"),
-        axis.text.y = element_text(colour="grey", size=rel(1.5), angle=0, hjust=.5, vjust=.5, face="plain"),
-        panel.background = element_blank(),
-        axis.line = element_line(colour = "grey")##,
-        )
-
-p4.2
-```
-
-## Statistical tests
-
-### Standard t-test
-
-Standard t-test comparing:
-
-- Population 1: Parents
-- Population 2: Offspring
-
-Assumptions:
-
-- Data distributed normal
-- Equal variance in the two groups (we can test this with the Levene's test later)
-
-Plotting the densities:
-
-```{r, echo=FALSE}
-q1 = ggplot(dat, aes(Cotyledons, fill=Generation))+geom_density(alpha=0.25)+
-  theme(
-        plot.title = element_text(hjust=0.5, size=rel(1.8)),
-        axis.title.x = element_text(size=rel(1.8)),
-        axis.title.y = element_text(size=rel(1.8), angle=90, vjust=0.5, hjust=0.5),
-        axis.text.x = element_text(colour="grey", size=rel(1.5), angle=0, hjust=.5, vjust=.5, face="plain"),
-        axis.text.y = element_text(colour="grey", size=rel(1.5), angle=0, hjust=.5, vjust=.5, face="plain"),
-        panel.background = element_blank(),
-        axis.line = element_line(colour = "grey")##,
-        )
-q1
-```
-
-Not really normal!
-
-```{r}
-x = dat$Cotyledons[dat$Generation == "P"]
-y = dat$Cotyledons[dat$Generation == "O"]
-```
-
-Variances look similar:
-```{r}
-var(x)
-var(y)
-```
-
-But let's do a Levene test:
-```{r}
-leveneTest(Cotyledons ~ Generation, dat, center=mean)
-```
-
-The p-value is less than 0.05, so we reject the null hypothesis of equal variances.
-
-The Levene test also assumes Normality. So, we can run the alternative Fligner-Killeen test with the same conclusion (equal variances):
-```{r}
-fligner.test(Cotyledons ~ Generation, dat)
-```
-
-Back to the t-test assumptions:
-
-- Normality: not met
-- Equality of variances: not met
-
-If we ignore the violations and we run a standard t-test, we get:
-```{r}
-t.test(x,y, var.equal=TRUE)
-```
-
-The p-value is less than 0.05, so we reject the null hypothesis of equality of means.
-This implies that we have evidence that the number of cotyledons is significantly larger in the offspring than in the parents.
-
-Our justification to do a standard t-test is that if the sample size is moderately large, the two-sample t-test is robust to non-normality due to the central limit theorem.
-
-We can also run the Welch version of t test that does not assume equal variances (this is the default in R):
-
-```{r}
-t.test(x,y)
-```
-
-### Wilcoxon-Mann-Whitney test
-
-Alternative to the t-test when normality is violated.
-This test does not test equality of means, but equality of distributions, so the conclusion could be deceiving (that is, two different distributions with the same mean will have a rejected null hypothesis using WMW test).
-
-```{r}
-wilcox.test(Cotyledons ~ Generation, dat)
-```
-
-The p-value less than 0.05 allows us to also reject the null hypothesis of equality of distributions which provides evidence that the number of cotyledons on the offspring group is significantly higher than in the parent group.
-
-# New data: `Polycot-data-2020.csv` focusing on chi-square test
-
-### Reading the data
-
-Reading in the data (had to remove manually the '#' sign from the column and save as csv file):
-```{r}
-dat = read.table("data/Polycot-data-2020.csv", sep=",",header=TRUE)
-str(dat)
-head(dat)
-summary(dat)
-dat = within(dat, Generation<-factor(Generation, levels=c("P","O")))
-```
-
-Creating categorical variable: dycots/polycots
-
-```{r}
-dat2 = data.frame(generation=dat$Generation)
-dat2$cots = "dicot"
-dat2$cots[dat$Cotyledons > 2] = "polycot"
-dat2 = within(dat2, cots<-factor(cots))
-summary(dat2)
-write.table(dat2,file="data/polycot-data-2020-categories.csv", row.names = FALSE, sep=",")
-
-dt = table(dat2$cots, dat2$generation)
-```
-
-### Data visualization
-```{r}
-library(graphics)
-mosaicplot(dt, shade = TRUE, las=2, main="Cotelydons")
-
-library(ggmosaic)
-ggplot(data = dat2) +
-  geom_mosaic(aes(x = product(cots), fill=generation))
-```
-
-### Data analysis
-```{r}
-dt
-ct = chisq.test(dt)
-
-obs = ct$observed
-exp = ct$expected
-
-df = rbind(obs[1,],exp[1,],obs[2,],exp[2,],colSums(obs))
-rownames(df) = c("dycot observed", "dycot expected", "polycot observed", "polycot expected", "total")
-
-r = rowSums(obs)
-ss = sum(r)
-totals = c(r[1],NA,r[2],NA,ss)
-cbind(df,totals)
-ct$statistic
-ct$p.value
-
-## expected values
-2437*2556/5585
-2437*3029/5585
-3148*2556/5585
-3148*3029/5585
-```

### 2022-03-19 ms/response2reviews.md
changes to notebooks
minor changes to responses to reviewers

@@
 We thank the reviewers for such helpful comments that have greatly improved the web apps and the manuscript.
+All of our changes are in the [review branch](https://github.com/crsl4/fast-stats/tree/review) which we will merge to master after approval by the reviewers and editorial team.
 
@@
 This is actually related to another point brought up by the reviewer (the notebooks).
-It was precisely the goal of the notebooks to be that reproducible script that anyone could run (but as the reviewer mentioned, they were not working properly).
+It was precisely the goal of the notebooks to be the reproducible script that anyone could run (but as the reviewer mentioned, they were not working properly).
 We have now updated the notebooks so that they are indeed reproducible so that anyone can run them and get the plots.

### 2022-03-31 notebooks/ecosystem-data-analysis.Rmd

@@
 ```{r}
-dat = read.table("data/ecosystem-data.csv", sep=",",header=TRUE)
+dat = read.table("../data/ecosystem-data.csv", sep=",",header=TRUE)
 str(dat)
